---
title: "Signals and backtests"
---

```{r, include = FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

Indicators describe the market; signals and backtests turn that description into decisions. This article covers the last three analysis methods. Two of them, from the [`Signals`](../reference/Signals.html) class, mark where one column of a data frame crosses another, which is how most simple trading rules are written. The third, from [`StrategyBacktests`](../reference/StrategyBacktests.html), runs a trading strategy over an instrument's past candles with the package's own backtesting engine and reports how it would have done. The Python library hands the same work to the Python package [backtesting.py](https://kernc.github.io/backtesting.py/); the R engine follows that package line by line, so a strategy gives the same trades and statistics in both.

The table below lists the three methods.

| Kind | Member | Description |
|---|---|---|
| method | [`is_cross_over`](#is_cross_over) | Marks the rows where one column rises above another |
| method | [`is_cross_under`](#is_cross_under) | Marks the rows where one column falls below another |
| method | [`run_backtest`](#run_backtest) | Runs a strategy over the instrument's candles and returns its statistics |

## Signals

The two signal methods are different from every other analysis method in one way: they do not fetch candles. They work on a data frame you already have, usually the result of an indicator method, so you choose the range once and then look for crossings in it. Neither method changes the data frame you pass in; each returns a copy with fresh row names and one new logical column.

The two tests compare the columns on the previous row and on this row. The table below states them exactly.

| Method | On the previous row | On this row | Column added |
|---|---|---|---|
| `is_cross_over` | first column at or below the second | first column above the second | `cross_over` |
| `is_cross_under` | first column at or above the second | first column below the second | `cross_under` |

The first row is never marked, because it has no previous row, and a row where either comparison meets a missing value is marked `FALSE`, as pandas does in the Python library. This is the textbook test, and it replaced the old project's version, which compared the previous first value with the current second value and could disagree when the second column moved sharply. On 400 days of INFY closes against a 20-candle simple moving average, the Python methods marked 16 crossovers and 17 crossunders on 2026-09-14, and on 2026-10-07 the R methods matched the Python ones on every row of the fixture candles they were checked against.

### is_cross_over

`is_cross_over(data, first_column, second_column)` marks each row where the first column rises above the second. A typical use is the close crossing above a moving average, or a fast average crossing above a slow one. Its reference entry is [`Signals$is_cross_over()`](../reference/Signals.html#method-Signals-is_cross_over).

#### Parameters

| Name | Type | Required | Default | Description |
|---|---|---|---|---|
| `data` | `data.frame` | Yes | | The data frame holding both columns. It is not changed. |
| `first_column` | character | Yes | | The name of the column that crosses |
| `second_column` | character | Yes | | The name of the column that is crossed |

#### Example

The example below was not run for this article, because it needs a live UBI. It finds the days in the last year on which RELIANCE closed above its 20-day simple moving average after closing at or below it the day before.

```{r}
reliance <- Equity$new(exchange = "nse", symbol = "RELIANCE")
frame <- reliance$simple_moving_average(window = 20, days = 365)

crossings <- reliance$is_cross_over(frame, "close", "sma_20")
columns <- c(
  "datetime",
  "close",
  "sma_20"
)
print(crossings[crossings$cross_over, columns])
```

#### Returns

A `data.frame` copy of `data` with fresh row names and an added logical `cross_over` column.

#### Errors

The table below lists the condition the method can signal.

| Condition | When |
|---|---|
| `KeyError` | `data` has no column named `first_column` or `second_column` |

### is_cross_under

`is_cross_under(data, first_column, second_column)` marks each row where the first column falls below the second. It is the mirror image of `is_cross_over`. Its reference entry is [`Signals$is_cross_under()`](../reference/Signals.html#method-Signals-is_cross_under).

#### Parameters

| Name | Type | Required | Default | Description |
|---|---|---|---|---|
| `data` | `data.frame` | Yes | | The data frame holding both columns. It is not changed. |
| `first_column` | character | Yes | | The name of the column that crosses |
| `second_column` | character | Yes | | The name of the column that is crossed |

#### Example

The example below, which was not run for this article, marks the days RELIANCE's MACD line fell below its signal line. The column names come from `moving_average_convergence_divergence()` with its default periods of 12, 26 and 9.

```{r}
frame <- reliance$moving_average_convergence_divergence(days = 365)
crossings <- reliance$is_cross_under(
  frame,
  "macd_12_26_9",
  "macd_12_26_9_signal"
)
columns <- c(
  "datetime",
  "close"
)
print(crossings[crossings$cross_under, columns])
```

#### Returns

A `data.frame` copy of `data` with fresh row names and an added logical `cross_under` column.

#### Errors

The table below lists the condition the method can signal.

| Condition | When |
|---|---|
| `KeyError` | `data` has no column named `first_column` or `second_column` |

## Backtests

A backtest replays a trading strategy over past candles as though it had been running then, and reports the trades it would have made and what they would have earned. `run_backtest()` does this with the package's backtesting engine, eight R6 classes in the files `R/assets_analysis_backtesting_*.R` that follow backtesting.py version 0.6.5, the package the Python library uses. You write the strategy as an R6 class that inherits [`BacktestStrategy`](../reference/BacktestStrategy.html); the method supplies the candles and runs it.

The sequence diagram below shows what happens inside one call.

<pre class="mermaid">
sequenceDiagram
    autonumber
    participant Y as Your code
    participant R as run_backtest
    participant U as UBI
    participant B as Backtest
    Y-&gt;&gt;R: strategy class, cash, commission, range
    R-&gt;&gt;U: GET /api/instruments/prices
    U--&gt;&gt;R: candles, or none
    alt no candles
        R--&gt;&gt;Y: NULL
    else candles
        R-&gt;&gt;R: keep datetime, open, high, low, close, volume
        R-&gt;&gt;R: rename to Open High Low Close Volume
        R-&gt;&gt;B: Backtest$new(candles, strategy, cash, commission, margin, ...)
        B-&gt;&gt;B: run(), which calls next_candle() once per candle
        B--&gt;&gt;R: statistics
        opt plot_filename given
            R-&gt;&gt;B: plot(results, filename)
        end
        R--&gt;&gt;Y: named list of statistics
    end
</pre>

The method selects the five candle columns by name and renames them to the capitalised names strategies use, so the extra columns UBI's candles carry, such as `oi` and `price_factor`, are dropped. The candles' `datetime` is kept with them, still in India time.

The engine follows backtesting.py's fill rules, which the table below summarises.

| Situation | Fill price |
|---|---|
| A market order | The next candle's open, or the previous close with `trade_on_close = TRUE`; a stop-loss or take-profit always uses the open |
| A stop order whose stop is reached | It becomes a market or limit order, and a market fill is no better than the stop |
| A limit order whose limit is reached | The limit, or the open, or the stop just reached, when that is better |
| A limit and a stop both reached on one candle | The limit is assumed to come first, so the order waits |

### Writing a strategy

A strategy subclass overrides two hooks, which the engine calls for you. The table below compares them, and the other parts of a strategy, with their backtesting.py counterparts, so that a strategy written for the Python library can be translated.

| backtesting.py | This package | Why it differs |
|---|---|---|
| `init()` | `initialize_strategy()` | `initialize()` is the R6 constructor |
| `next()` | `next_candle()` | `next` is a reserved word in R |
| `self.I(func, *args)` assigned to an attribute | `self$indicator(name, values)`, read back as `self$indicators$name` | An R6 object is locked once it is built and cannot gain fields, so indicators are stored by name; you compute `values` yourself, such as `talib::SMA(self$data$Close, timePeriod = 10)` |
| `self.data.Close[-1]` | `tail(self$data$Close, 1)` | R has no negative indexing from the end |
| `if not self.position:` | `if (self$position$size == 0)` | An R object has no truth value |
| `backtesting.lib.crossover(a, b)` | `self$crossover(a, b)` | Behaviour lives on the class that uses it |

Inside `next_candle()`, `self$data` and `self$indicators` show only the candles up to the current one, exactly as backtesting.py's arrays do, so a strategy cannot look ahead. `self$buy()` and `self$sell()` take a `size` that is either a fraction of the equity strictly between 0 and 1 or a whole number of at least 1, and they default to nearly all of the equity, as in backtesting.py. Indicators must be numeric vectors; backtesting.py's two-dimensional indicators and its plotting options are not ported.

### run_backtest

`run_backtest(strategy, cash = 10000, commission = 0.0, margin = 1.0, trade_on_close = FALSE, hedging = FALSE, exclusive_orders = FALSE, plot_filename = NULL, interval = "day", from_date = NULL, to_date = NULL, days = NULL, adjusted = TRUE)` sends `GET /api/instruments/prices`. Its reference entry is [`StrategyBacktests$run_backtest()`](../reference/StrategyBacktests.html#method-StrategyBacktests-run_backtest).

This method runs a strategy over the instrument's candles in a range and returns the backtest's statistics. It sends no orders; it only reads candles. When `plot_filename` is given, it also writes a plot of the run to that HTML file without opening it.

#### Parameters

| Name | Type | Required | Default | Description |
|---|---|---|---|---|
| `strategy` | R6 class generator | Yes | | The strategy class to run, a subclass of `BacktestStrategy`, not an object built from it |
| `cash` | numeric | No | `10000` | The starting cash |
| `commission` | numeric | No | `0.0` | The commission on each trade, as a fraction of its value |
| `margin` | numeric | No | `1.0` | The margin required, as a fraction, where 1.0 means no leverage |
| `trade_on_close` | logical | No | `FALSE` | `TRUE` to fill market orders at the current candle's close rather than the next candle's open |
| `hedging` | logical | No | `FALSE` | `TRUE` to allow long and short trades at the same time |
| `exclusive_orders` | logical | No | `FALSE` | `TRUE` to close the open trade whenever a new order is placed |
| `plot_filename` | character | No | `NULL` | The path of an HTML file to write the plot to, or `NULL` for no plot |
| `interval`, `from_date`, `to_date`, `days`, `adjusted` | | No | as in `prices()` | The candle range, as described in [Analysis](analysis.html#the-common-arguments) |

#### Example

The example below defines the classic moving average crossover strategy, which buys when a 10-candle average crosses above a 30-candle one and sells on the reverse, and runs it over two years of RELIANCE with 100,000 rupees. It was not run for this article, because it needs a live UBI. The same strategy over 730 days of INFY reported 9 trades when the Python method was checked on 2026-09-14.

```{r}
SmaCross <- R6::R6Class(
  "SmaCross",
  inherit = BacktestStrategy,
  public = list(
    fast = 10,
    slow = 30,

    initialize_strategy = function() {
      self$indicator(
        "fast_average",
        talib::SMA(self$data$Close, timePeriod = self$fast)
      )
      self$indicator(
        "slow_average",
        talib::SMA(self$data$Close, timePeriod = self$slow)
      )
    },

    next_candle = function() {
      fast_average <- self$indicators$fast_average
      slow_average <- self$indicators$slow_average
      if (self$crossover(fast_average, slow_average)) {
        self$position$close()
        self$buy()
      } else if (self$crossover(slow_average, fast_average)) {
        self$position$close()
        self$sell()
      }
    }
  )
)

reliance <- Equity$new(exchange = "nse", symbol = "RELIANCE")
statistics <- reliance$run_backtest(
  SmaCross,
  cash = 100000,
  commission = 0.001,
  days = 730,
  plot_filename = "sma_cross.html"
)
labels <- c(
  "Return [%]",
  "# Trades",
  "Win Rate [%]",
  "Max. Drawdown [%]"
)
str(statistics[labels])
```

The strategy's hooks are `initialize_strategy()` and `next_candle()`, the R names of backtesting.py's `init` and `next`. The statistics labels in the last lines are backtesting.py's own, kept unchanged so that results can be compared with the Python library; backtesting.py's documentation lists them all.

#### Returns

A named list of the backtest's statistics, with the same names in the same order as the pandas Series backtesting.py returns, or `NULL` when UBI has no candles for the range. `Start` and `End` are `POSIXct` values, and `Duration` and the four duration statistics are `difftime` values in days. The trades themselves are a data frame under `_trades`, with `EntryTime` and `ExitTime` columns and candle numbers in `EntryBar` and `ExitBar` that count from 1, so each is one more than in Python. The equity curve is a data frame under `_equity_curve`, with a `datetime` column first. Trades still open at the end count in the equity curve but not in the trade statistics, as in backtesting.py, and a warning says so.

#### Errors

The table below lists the condition the method can signal.

| Condition | When |
|---|---|
| `UnifiedBrokerInterfaceError` | UBI refused the candle request or could not be reached |

::: {.alert .alert-info}
**Under the hood.**

The old project always called backtesting.py's `plot()`, which wrote an HTML file named after the strategy into the current directory and opened a browser tab on every run. In both the Python library and this package the plot is optional and never opens anything by itself, and it shows the run just made rather than running the strategy a second time. The R plot is not backtesting.py's interactive Bokeh chart, which has no R counterpart. It is one self-contained HTML page with no scripts and no network requests, holding a line of the closes with a green dot at each closed trade's entry and a red dot at its exit, a line of the equity, the statistics as a table and the closed trades as a table.
:::

On 2026-10-07 fourteen backtests on the parity fixtures, covering commissions, `trade_on_close`, `exclusive_orders`, a margin of 0.5, stop-loss and take-profit orders, stop and limit entries with partial closes, hedging and 5-minute candles, gave the same trades on the same candles at the same prices in R as in backtesting.py. The largest relative difference in any statistic was 2.7e-14. backtesting.py's bid-ask spread, commission given as a function, strategy parameters passed to `run()` and `optimize()` are not ported, because `run_backtest()` never uses them.

A backtest is only as good as its candles. Run it on an instrument that has them, which today means shares, equity indices, exchange traded funds and commodity derivatives, as [Which instruments have candles](analysis.html#which-instruments-have-candles) explains. Shares and funds are adjusted for splits and bonuses by default, which is what a backtest over several years needs.
