---
title: "Indicators"
---

```{r, include = FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

Technical indicators are calculations over a run of candles that traders read as signs of trend, momentum or volatility. This article lists the 52 indicator methods every instrument inherits, in five groups that follow TA-Lib's own grouping. Each method fetches the instrument's candles through [`prices()`](../reference/Instrument.html#method-Instrument-prices), runs one TA-Lib function over them, through the R `talib` package or a base R version of the same calculation, and returns the candles with one or more columns added, or `NULL` when UBI has no candles for the range.

The table below shows the five groups and what each is for.

| Group | Class | Methods | What it measures |
|---|---|--:|---|
| [Overlap studies](#overlap-studies) | [`OverlapStudies`](../reference/OverlapStudies.html) | 12 | Averages and bands drawn on the same scale as the price |
| [Momentum indicators](#momentum-indicators) | [`MomentumIndicators`](../reference/MomentumIndicators.html) | 28 | How fast and how strongly the price is moving |
| [Volume indicators](#volume-indicators) | [`VolumeIndicators`](../reference/VolumeIndicators.html) | 3 | Whether volume confirms the price's moves |
| [Volatility indicators](#volatility-indicators) | [`VolatilityIndicators`](../reference/VolatilityIndicators.html) | 3 | How far the price ranges from candle to candle |
| [Cycle indicators](#cycle-indicators) | [`CycleIndicators`](../reference/CycleIndicators.html) | 6 | Repeating cycles, found with the Hilbert transform |

The flowchart below is a rough guide to which group answers which question, with one commonly used method from each.

<pre class="mermaid">
flowchart LR
    Q{"What do you&lt;br/&gt;want to know?"}
    Q --&gt;|"Which way is the trend?"| O["Overlap studies&lt;br/&gt;simple_moving_average"]
    Q --&gt;|"Is the move running out of steam?"| M["Momentum indicators&lt;br/&gt;relative_strength_index"]
    Q --&gt;|"Is volume behind the move?"| V["Volume indicators&lt;br/&gt;on_balance_volume"]
    Q --&gt;|"How far does it swing?"| A["Volatility indicators&lt;br/&gt;average_true_range"]
    Q --&gt;|"Is it trending or cycling?"| C["Cycle indicators&lt;br/&gt;hilbert_transform_trend_mode"]
</pre>

## Arguments every method shares

Every method in this article also takes the five range arguments of `prices()`, which are `interval`, `from_date`, `to_date`, `days` and `adjusted`. They are left out of the tables below to keep them readable, and [Analysis](analysis.html#the-common-arguments) describes them. Pass every argument by name, as in `simple_moving_average(window = 20, days = 365)`, because the argument order is not the same from method to method.

The tables use the package's argument names, which spell out TA-Lib's. The table below maps the ones that differ most, both to the names Python's `talib` uses and to the names of the R `talib` package, which the methods pass them to.

| This package | Python's `talib` | R's `talib` package |
|---|---|---|
| `window` | `timeperiod` | `timePeriod` |
| `fast_period`, `slow_period`, `signal_period` | `fastperiod`, `slowperiod`, `signalperiod` | `fastPeriod`, `slowPeriod`, `signalPeriod` |
| `moving_average_type` | `matype` | `maType` |
| `fast_k_period`, `slow_k_period`, `slow_d_period`, `fast_d_period` | `fastk_period`, `slowk_period`, `slowd_period`, `fastd_period` | `fastKPeriod`, `slowKPeriod`, `slowDPeriod`, `fastDPeriod` |
| `standard_deviations_up`, `standard_deviations_down` | `nbdevup`, `nbdevdn` | `deviationsUp`, `deviationsDown` |
| `volume_factor` | `vfactor` | `volumeFactor` |
| `fast_limit`, `slow_limit` | `fastlimit`, `slowlimit` | `fastLimit`, `slowLimit` |

Several methods take a moving average type, an integer that chooses which kind of average TA-Lib uses inside the calculation. The values are TA-Lib's own, as the table below lists; `0`, a simple moving average, is the default everywhere. The `talib` package passes these integers straight to the C library, so each one selects the same average in R as in Python. The `talib` package's own `APO` and `PPO` default to 1 rather than 0, but the methods always pass the argument, so the package's defaults never apply.

| Value | Moving average |
|--:|---|
| 0 | Simple |
| 1 | Exponential |
| 2 | Weighted |
| 3 | Double exponential |
| 4 | Triple exponential |
| 5 | Triangular |
| 6 | Kaufman adaptive |
| 7 | MESA adaptive |
| 8 | Triple generalised double exponential (T3) |

## The methods

The sections below hold one table per group, in the same order as the table at the top of the article. The "Columns added" column gives each new column's name, in which a placeholder such as `<window>` is replaced by the argument's value, so `relative_strength_index(window = 14)` adds `rsi_14`. The "TA-Lib function" column names the TA-Lib function each method computes; in R all but two are called through the `talib` package under the same name, such as `talib::SMA`, and the two written in base R, `ROC` and `ROCP`, are described below the momentum table.

### Overlap studies

Overlap studies are drawn on the same scale as the price, so they can be plotted over the candles. Moving averages smooth the price, Bollinger bands put an envelope around it, and the parabolic SAR marks a trailing stop level.

| Method | What it adds | Own arguments and defaults | TA-Lib function | Columns added |
|---|---|---|---|---|
| `simple_moving_average` | Adds the simple moving average of one candle column | `window = 10`, `column = "close"` | `SMA` | `sma_<window>` |
| `exponential_moving_average` | Adds the exponential moving average of one candle column | `window = 10`, `column = "close"` | `EMA` | `ema_<window>` |
| `bollinger_bands` | Adds the upper, middle and lower Bollinger bands of one candle column | `window = 10`, `standard_deviations_up = 2`, `standard_deviations_down = 2`, `column = "close"` | `BBANDS` | `bb_upper_<window>`, `bb_middle_<window>`, `bb_lower_<window>` |
| `weighted_moving_average` | Adds the weighted moving average of one candle column | `window = 10`, `column = "close"` | `WMA` | `wma_<window>` |
| `double_exponential_moving_average` | Adds the double exponential moving average of one candle column | `window = 10`, `column = "close"` | `DEMA` | `dema_<window>` |
| `triple_exponential_moving_average` | Adds Tillson's T3 triple exponential moving average of one candle column | `window = 10`, `volume_factor = 0.7`, `column = "close"` | `T3` | `t3_<window>` |
| `kaufman_adaptive_moving_average` | Adds the Kaufman adaptive moving average of one candle column | `window = 10`, `column = "close"` | `KAMA` | `kama_<window>` |
| `mesa_adaptive_moving_average` | Adds the MESA adaptive moving average and its following average of one candle column | `fast_limit = 0.5`, `slow_limit = 0.05`, `column = "close"` | `MAMA` | `mama`, `fama` |
| `triangular_moving_average` | Adds the triangular moving average of one candle column | `window = 10`, `column = "close"` | `TRIMA` | `trima_<window>` |
| `parabolic_sar` | Adds the parabolic stop and reverse from the high and low columns | `acceleration = 0.02`, `maximum = 0.2` | `SAR` | `psar` |
| `mid_point` | Adds the midpoint of the highest and lowest value of one candle column over each window | `window = 10`, `column = "close"` | `MIDPOINT` | `mid_point_<window>` |
| `middle_price` | Adds the midpoint of the highest high and lowest low over each window | `window = 10` | `MIDPRICE` | `middle_price_<window>` |

`triple_exponential_moving_average` calls TA-Lib's `T3`, Tillson's T3, rather than `TEMA`, which is the indicator most charting tools call the triple exponential moving average. The name was kept from the old project, and its column is `t3_<window>`.

### Momentum indicators

Momentum indicators measure how fast the price is moving and whether that speed is growing or fading. Most of them are oscillators that move between fixed bounds, such as the relative strength index between 0 and 100.

| Method | What it adds | Own arguments and defaults | TA-Lib function | Columns added |
|---|---|---|---|---|
| `moving_average_convergence_divergence` | Adds the moving average convergence divergence line, its signal line and their difference | `fast_period = 12`, `slow_period = 26`, `signal_period = 9`, `column = "close"` | `MACD` | `macd_<fast>_<slow>_<signal>`, `macd_<fast>_<slow>_<signal>_signal`, `macd_<fast>_<slow>_<signal>_hist` |
| `average_directional_movement_index` | Adds the average directional movement index | `window = 14` | `ADX` | `adx_<window>` |
| `momentum` | Adds the momentum of one candle column, its change over each window | `window = 14`, `column = "close"` | `MOM` | `momentum_<window>` |
| `commodity_channel_index` | Adds the commodity channel index | `window = 14` | `CCI` | `cci_<window>` |
| `average_directional_movement_index_rating` | Adds the average directional movement index rating | `window = 10` | `ADXR` | `adxr_<window>` |
| `absolute_price_oscillator` | Adds the absolute price oscillator, the difference between a fast and a slow moving average | `fast_period = 12`, `slow_period = 26`, `moving_average_type = 0`, `column = "close"` | `APO` | `apo_<fast>_<slow>` |
| `aroon` | Adds the Aroon down and Aroon up lines | `window = 10` | `AROON` | `aroon_down_<window>`, `aroon_up_<window>` |
| `aroon_oscillator` | Adds the Aroon oscillator, Aroon up minus Aroon down | `window = 10` | `AROONOSC` | `aroon_osc_<window>` |
| `balance_of_power` | Adds the balance of power | none | `BOP` | `bop` |
| `chande_momentum_oscillator` | Adds the Chande momentum oscillator of one candle column | `window = 10`, `column = "close"` | `CMO` | `cmo_<window>` |
| `directional_movement_index` | Adds the directional movement index | `window = 10` | `DX` | `dx_<window>` |
| `moving_average_convergence_divergence_extended` | Adds the moving average convergence divergence with a chosen moving average type for each of its three averages | `fast_period = 12`, `fast_moving_average_type = 0`, `slow_period = 26`, `slow_moving_average_type = 0`, `signal_period = 9`, `signal_moving_average_type = 0`, `column = "close"` | `MACDEXT` | `macd_<fast>_<slow>_<signal>`, `macd_signal_<fast>_<slow>_<signal>`, `macd_hist_<fast>_<slow>_<signal>` |
| `money_flow_index` | Adds the money flow index, a relative strength index weighted by volume | `window = 14` | `MFI` | `mfi_<window>` |
| `minus_directional_indicator` | Adds the minus directional indicator | `window = 14` | `MINUS_DI` | `minus_di_<window>` |
| `minus_directional_movement` | Adds the minus directional movement | `window = 14` | `MINUS_DM` | `minus_dm_<window>` |
| `plus_directional_indicator` | Adds the plus directional indicator | `window = 14` | `PLUS_DI` | `plus_di_<window>` |
| `plus_directional_movement` | Adds the plus directional movement | `window = 14` | `PLUS_DM` | `plus_dm_<window>` |
| `percentage_price_oscillator` | Adds the percentage price oscillator, the gap between a fast and a slow moving average as a percentage | `fast_period = 12`, `slow_period = 26`, `moving_average_type = 0`, `column = "close"` | `PPO` | `ppo<fast>_<slow>` |
| `rate_of_change` | Adds the rate of change of one candle column as a percentage | `window = 14`, `column = "close"` | `ROC` | `roc_<window>` |
| `rate_of_change_percent` | Adds the rate of change of one candle column as a fraction | `window = 14`, `column = "close"` | `ROCP` | `rocp_<window>` |
| `rate_of_change_ratio` | Adds the rate of change of one candle column as a ratio | `window = 14`, `column = "close"` | `ROCR` | `rocr_<window>` |
| `relative_strength_index` | Adds the relative strength index of one candle column | `window = 14`, `column = "close"` | `RSI` | `rsi_<window>` |
| `stochastic_oscillator` | Adds the slow stochastic oscillator's %K and %D lines | `fast_k_period = 5`, `slow_k_period = 3`, `slow_k_moving_average_type = 0`, `slow_d_period = 3`, `slow_d_moving_average_type = 0` | `STOCH` | `slowk_<slow_k_period>`, `slowd_<slow_d_period>` |
| `stochastic_fast_oscillator` | Adds the fast stochastic oscillator's %K and %D lines | `fast_k_period = 5`, `fast_d_period = 3`, `fast_d_moving_average_type = 0` | `STOCHF` | `stochf_fastk<fast_k_period>`, `stochf_fastd<fast_d_period>` |
| `stochastic_relative_strength_index` | Adds the stochastic relative strength index's %K and %D lines for one candle column | `window = 14`, `fast_k_period = 5`, `fast_d_period = 3`, `fast_d_moving_average_type = 0`, `column = "close"` | `STOCHRSI` | `stochrsi_fastk<fast_k_period>`, `stochrsi_fastd<fast_d_period>` |
| `trix` | Adds TRIX, the rate of change of a triple smoothed exponential moving average of one candle column | `window = 15`, `column = "close"` | `TRIX` | `trix_<window>` |
| `ultimate_oscillator` | Adds the ultimate oscillator, which blends buying pressure over three windows | `fast_period = 7`, `slow_period = 14`, `signal_period = 28` | `ULTOSC` | `ultosc_<fast>_<slow>_<signal>` |
| `williams_percent_r` | Adds Williams %R | `window = 14` | `WILLR` | `willr_<window>` |

A few labels are irregular, and they are kept as the old project wrote them so that existing code reading the data frames still works, in Python and in R alike. The two MACD methods name their signal and histogram columns differently, `ppo`, `stochf_` and `stochrsi_` labels have no underscore before their numbers, and `ultimate_oscillator` calls its short, middle and long windows `fast_period`, `slow_period` and `signal_period`. The directional movement defaults also differ: `average_directional_movement_index` uses 14 candles, while `average_directional_movement_index_rating` and `directional_movement_index` use 10.

`stochastic_relative_strength_index` takes a separate `window` for the length of the underlying RSI, 14 by default, and passes `fast_k_period` to TA-Lib's fast %K period, which the `talib` package calls `fastKPeriod`. With its defaults it matches Python's `talib.STOCHRSI` with TA-Lib's own defaults.

`rate_of_change` and `rate_of_change_percent` are written in base R, because the `talib` package has `ROCR` but no `ROC` or `ROCP`. They follow TA-Lib's C definitions, `(value / earlier - 1) * 100` and `(value - earlier) / earlier`, in the same order of operations, leave the first `window` rows empty, and give 0 when the earlier value is 0, so their values are identical to Python's. Every other momentum method calls the `talib` package.

::: {.alert .alert-warning}
**One known difference from the Python library.**

`stochastic_relative_strength_index` with `fast_d_moving_average_type = 6`, a Kaufman adaptive average smoothing the %D line, did not match the Python library on the fixture candles it was checked against on 2026-10-07. The %K line matched exactly, but %D differed on 61 of 400 rows, by up to 12.9 points. The cause is a change in TA-Lib itself: the `talib` package bundles TA-Lib C 0.8.1, while Python's `talib` links TA-Lib C 0.6.4, and when %K is almost flat, such as pinned at 100, the Kaufman average divides two numbers that are only rounding residue, which version 0.8.1 calculates differently on purpose. A one-bit difference can then send the two versions to opposite ends of the smoothing range, and the gap closes again over the following rows. `stochastic_oscillator` and `stochastic_fast_oscillator` with the same moving average type could in principle hit the same difference, although the fixture did not trigger it. Every other momentum case matched the Python library within a relative difference of 5e-11.
:::

### Volume indicators

Volume indicators combine price and traded volume, to show whether buying or selling pressure is behind a move. They need a `volume` column, so they are only meaningful for instruments whose candles carry real volume.

| Method | What it adds | Own arguments and defaults | TA-Lib function | Columns added |
|---|---|---|---|---|
| `chaikin_accumulation_distribution_line` | Adds the Chaikin accumulation distribution line | none | `AD` | `chaikin_ad` |
| `chaikin_accumulation_distribution_oscillator` | Adds the Chaikin accumulation distribution oscillator | `fast_period = 3`, `slow_period = 10` | `ADOSC` | `chaikin_adosc<fast>_<slow>` |
| `on_balance_volume` | Adds the on balance volume, measured against one candle column | `column = "close"` | `OBV` | `obv` |

### Volatility indicators

Volatility indicators measure how far the price ranges. The true range of a candle is the largest of its high minus its low and the gaps from the previous close, and the average true range smooths it over a window.

| Method | What it adds | Own arguments and defaults | TA-Lib function | Columns added |
|---|---|---|---|---|
| `average_true_range` | Adds the average true range | `window = 14` | `ATR` | `atr_<window>` |
| `normalized_average_true_range` | Adds the average true range as a percentage of the close | `window = 14` | `NATR` | `natr<window>` |
| `true_range` | Adds each candle's true range | none | `TRANGE` | `tr` |

The normalised average true range's column is `natr<window>`, without the underscore that `atr_<window>` has.

### Cycle indicators

Cycle indicators use John Ehlers' Hilbert transform to find a repeating cycle in the price and say whether the market is trending or cycling. They take no window, because the transform chooses its own.

| Method | What it adds | Own arguments and defaults | TA-Lib function | Columns added |
|---|---|---|---|---|
| `hilbert_transform_dominant_cycle_period` | Adds the Hilbert transform dominant cycle period of one candle column | `column = "close"` | `HT_DCPERIOD` | `ht_dcperiod` |
| `hilbert_transform_dominant_cycle_phase` | Adds the Hilbert transform dominant cycle phase of one candle column | `column = "close"` | `HT_DCPHASE` | `ht_dcphase` |
| `hilbert_transform_phasor_components` | Adds the Hilbert transform in-phase and quadrature phasor components of one candle column | `column = "close"` | `HT_PHASOR` | `inphase`, `quadrature` |
| `hilbert_transform_sine_wave` | Adds the Hilbert transform sine wave and lead sine wave of one candle column | `column = "close"` | `HT_SINE` | `sine`, `lead_sine` |
| `hilbert_transform_trend_mode` | Adds the Hilbert transform trend mode of one candle column, 1 in a trend and 0 in a cycle | `column = "close"` | `HT_TRENDMODE` | `ht_trendmode` |
| `hilbert_transform_trend_line` | Adds the Hilbert transform instantaneous trend line of one candle column | `column = "close"` | `HT_TRENDLINE` | `ht_trendline` |

These need a long run of candles before their first value. TA-Lib reports a lookback of 32 candles for `HT_DCPERIOD` and `HT_PHASOR`, and 63 for the other four, so a short range gives columns that are mostly or entirely empty.

## A worked example

The example below computes a 14-day relative strength index for the nearest MCX gold future and prints three of the candles' columns.

```{r}
expiries <- CommodityFutures$expiries(
  exchange = "mcx",
  underlying_symbol = "GOLD"
)
gold <- CommodityFutures$new(
  exchange = "mcx",
  underlying_symbol = "GOLD",
  expiry_date = expiries[[1]]
)

frame <- gold$relative_strength_index(window = 14, days = 90)
columns <- c(
  "datetime",
  "close",
  "rsi_14"
)
print(frame[, columns])
```

When the Python library ran the same calculation against a local UBI on 2026-09-20, the call returned 61 rows. On the three days that check kept, 14 to 16 September 2026, gold closed at 151,230, 150,809 and 152,470 rupees, and its RSI read 44.26, 43.35 and 47.89. An RSI below 30 is conventionally read as oversold and one above 70 as overbought, so gold's readings in the mid-forties that week were neutral.

Because every method returns the candles, you can combine two indicators by joining their new columns. The next example, which was not run for this article, keeps the days on which a share closed below its lower Bollinger band while its RSI was under 30. Each method fetches the candles again, but both requests ask for the same range, so the rows line up.

```{r}
reliance <- Equity$new(exchange = "nse", symbol = "RELIANCE")

bands <- reliance$bollinger_bands(window = 20, days = 365)
strength <- reliance$relative_strength_index(window = 14, days = 365)

strength_columns <- c(
  "datetime",
  "rsi_14"
)
frame <- merge(bands, strength[, strength_columns], by = "datetime")
below_band <- frame$close < frame$bb_lower_20
weak_strength <- frame$rsi_14 < 30
oversold <- frame[which(below_band & weak_strength), ]
columns <- c(
  "datetime",
  "close",
  "bb_lower_20",
  "rsi_14"
)
print(oversold[, columns])
```

Remember that either call can return `NULL`, for an instrument or range with no candles, so real code checks for that with `is.null()` before merging. The example uses `which()` so that rows where either column is still `NA`, during the indicators' warm-up, are left out rather than turned into rows of `NA`.
