---
title: "Statistics and transforms"
---

```{r, include = FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

This article covers the 76 analysis methods that are not indicators or patterns. They fall into five classes: summary statistics of a range of candles, rolling statistic functions such as beta and linear regression, arithmetic between candle columns, mathematical transforms of one column, and the four price transforms that combine a candle's open, high, low and close into one figure.

The table below shows the five groups and how their results differ, which matters more here than in the other articles: the summary statistics return a single number, while the other four return the candles with a column added.

| Group | Class | Methods | Returns |
|---|---|--:|---|
| [Price statistics](#price-statistics) | [`PriceStatistics`](../reference/PriceStatistics.html) | 39 | Mostly a single number for the whole range |
| [Statistic functions](#statistic-functions) | [`StatisticFunctions`](../reference/StatisticFunctions.html) | 8 | The candles with a rolling column added |
| [Math operators](#math-operators) | [`MathOperators`](../reference/MathOperators.html) | 10 | The candles with a column added |
| [Math transforms](#math-transforms) | [`MathTransforms`](../reference/MathTransforms.html) | 15 | The candles with a column added |
| [Price transforms](#price-transforms) | [`PriceTransforms`](../reference/PriceTransforms.html) | 4 | The candles with a column added |

Every method in this article also takes the five range arguments of [`prices()`](../reference/Instrument.html#method-Instrument-prices), which are `interval`, `from_date`, `to_date`, `days` and `adjusted`, and returns `NULL` when UBI has no candles for the range. [Analysis](analysis.html#the-common-arguments) describes those arguments, and the tables below leave them out.

## Price statistics

`PriceStatistics` answers questions about a whole range at once, such as "what was the highest high this year" or "how skewed were the daily returns". Its 39 methods form three families, each built on the one before, as the diagram below shows.

<pre class="mermaid">
flowchart LR
    P["prices()"] --&gt; A["price_high, price_low&lt;br/&gt;price_mean ... price_histogram&lt;br/&gt;12 methods"]
    P --&gt; V["volumes()"]
    V --&gt; B["volume_total, volume_high&lt;br/&gt;... volume_histogram&lt;br/&gt;13 methods"]
    P --&gt; R["returns(column)"]
    R --&gt; C["returns_high, returns_low&lt;br/&gt;... returns_summary&lt;br/&gt;12 methods"]
</pre>

`volumes` and `returns` are themselves methods, and they return a narrowed data frame of `exchange`, `segment`, `datetime`, `interval` and one value column. `returns` divides each value by the one before it and subtracts 1, which is what pandas' `pct_change` does in the Python library, so its value is the fractional change from one candle to the next, 0.01 for a one per cent rise, and its first row is always `NA`. A missing value in the column gives `NA` on both sides of it, as pandas 3 does. `price_high` and `price_low` always read the `high` and `low` columns and take no `column` argument.

The table below lists all 39 methods.

| Method | What it does | Own arguments and defaults | Returns |
|---|---|---|---|
| `price_high` | Finds the highest high in the range | none | The highest high as a number |
| `price_low` | Finds the lowest low in the range | none | The lowest low as a number |
| `price_mean` | Finds the mean of one candle column in the range | `column = "close"` | The mean as a number |
| `price_median` | Finds the median of one candle column in the range | `column = "close"` | The median as a number |
| `price_standard_deviation` | Finds the standard deviation of one candle column in the range | `column = "close"` | The standard deviation as a number |
| `price_variance` | Finds the variance of one candle column in the range | `column = "close"` | The variance as a number |
| `price_mean_absolute_deviation` | Finds the mean absolute deviation of one candle column from its mean in the range | `column = "close"` | The mean absolute deviation as a number |
| `price_skewness` | Finds the skewness of one candle column in the range | `column = "close"` | The skewness as a number |
| `price_kurtosis` | Finds the kurtosis of one candle column in the range | `column = "close"` | The kurtosis as a number |
| `price_quantile` | Finds a quantile of one candle column in the range | `quantile = 0.5`, `column = "close"` | The quantile as a number |
| `price_summary` | Summarises one candle column in the range with count, mean, spread and quartiles | `column = "close"` | A named numeric vector of summary statistics |
| `price_histogram` | Draws a histogram of one candle column in the range with `graphics::hist()` | `bins = 50`, `column = "close"` | The `histogram` object `graphics::hist()` returns |
| `volumes` | Fetches the traded volume of each candle in the range | none | A data frame with `exchange`, `segment`, `datetime`, `interval` and `volume` columns |
| `volume_total` | Finds the total volume traded in the range | none | The total volume as a number |
| `volume_high` | Finds the highest volume of any candle in the range | none | The highest volume as a number |
| `volume_low` | Finds the lowest volume of any candle in the range | none | The lowest volume as a number |
| `volume_mean` | Finds the mean volume per candle in the range | none | The mean volume as a number |
| `volume_median` | Finds the median volume per candle in the range | none | The median volume as a number |
| `volume_standard_deviation` | Finds the standard deviation of volume per candle in the range | none | The standard deviation as a number |
| `volume_variance` | Finds the variance of volume per candle in the range | none | The variance as a number |
| `volume_mean_absolute_deviation` | Finds the mean absolute deviation of volume per candle from its mean in the range | none | The mean absolute deviation as a number |
| `volume_kurtosis` | Finds the kurtosis of volume per candle in the range | none | The kurtosis as a number |
| `volume_skewness` | Finds the skewness of volume per candle in the range | none | The skewness as a number |
| `volume_quantile` | Finds a quantile of volume per candle in the range | `quantile = 0.5` | The quantile as a number |
| `volume_summary` | Summarises volume per candle in the range with count, mean, spread and quartiles | none | A named numeric vector of summary statistics |
| `volume_histogram` | Draws a histogram of volume per candle in the range with `graphics::hist()` | `bins = 50` | The `histogram` object `graphics::hist()` returns |
| `returns` | Calculates the fractional change of one candle column from each candle to the next | `column = "close"` | A data frame with `exchange`, `segment`, `datetime`, `interval` and `returns` columns |
| `returns_high` | Finds the highest return of one candle column in the range | `column = "close"` | The highest return as a number |
| `returns_low` | Finds the lowest return of one candle column in the range | `column = "close"` | The lowest return as a number |
| `returns_mean` | Finds the mean return of one candle column in the range | `column = "close"` | The mean return as a number |
| `returns_median` | Finds the median return of one candle column in the range | `column = "close"` | The median return as a number |
| `returns_standard_deviation` | Finds the standard deviation of returns of one candle column in the range | `column = "close"` | The standard deviation as a number |
| `returns_variance` | Finds the variance of returns of one candle column in the range | `column = "close"` | The variance as a number |
| `returns_mean_absolute_deviation` | Finds the mean absolute deviation of returns of one candle column from their mean in the range | `column = "close"` | The mean absolute deviation as a number |
| `returns_skewness` | Finds the skewness of returns of one candle column in the range | `column = "close"` | The skewness as a number |
| `returns_kurtosis` | Finds the kurtosis of returns of one candle column in the range | `column = "close"` | The kurtosis as a number |
| `returns_quantile` | Finds a quantile of returns of one candle column in the range | `quantile = 0.5`, `column = "close"` | The quantile as a number |
| `returns_histogram` | Draws a histogram of returns of one candle column in the range with `graphics::hist()` | `bins = 50`, `column = "close"` | The `histogram` object `graphics::hist()` returns |
| `returns_summary` | Summarises returns of one candle column in the range with count, mean, spread and quartiles | `column = "close"` | A named numeric vector of summary statistics |

Each method follows the matching pandas reduction exactly, because base R's defaults differ from pandas' in several places. Missing values are skipped, as pandas skips them; the standard deviation and variance are sample figures, with n - 1 in the denominator; skewness and kurtosis use pandas' adjusted formulas, giving 0 for constant values and `NA` when there are too few values; quantiles use linear interpolation, `stats::quantile(type = 7)`; and volumes are summed as doubles, because an R integer sum overflows above about 2.1 billion. On 2026-10-07 every method was compared with the Python library on four fixtures, and the largest relative difference was about 1e-13.

The three summary methods return a named numeric vector with the names `count`, `mean`, `std`, `min`, `25%`, `50%`, `75%` and `max`, where the Python library returns a pandas Series with that index. `summary[["mean"]]` reads the same figure as Python's `summary["mean"]`.

::: {.alert .alert-info}
**The histogram methods draw with base R graphics.**

`price_histogram`, `volume_histogram` and `returns_histogram` draw with `graphics::hist()` on the current graphics device and return its `histogram` object invisibly, where the Python library draws with matplotlib and returns the matplotlib `Axes`. They need no extra package. The bins copy NumPy's rules, so the bar heights and edges are the same as the Python library's: a bar's edges are `histogram$breaks[[bar]]` and `histogram$breaks[[bar + 1]]`, and its height is `histogram$counts[[bar]]`. When there is no screen, open a file device such as `grDevices::png()` first, and close it with `grDevices::dev.off()` afterwards.
:::

The example below, which was not run for this article, summarises a year of RELIANCE's daily returns.

```{r}
reliance <- Equity$new(exchange = "nse", symbol = "RELIANCE")
print(reliance$returns_standard_deviation(days = 365))
print(reliance$returns_summary(days = 365))
```

## Statistic functions

`StatisticFunctions` holds TA-Lib's eight rolling statistic functions. Each works over a moving window and adds a column with one value per candle. The table below lists them.

| Method | What it adds | Own arguments and defaults | TA-Lib function | Columns added |
|---|---|---|---|---|
| `beta` | Adds the rolling beta of the instrument against a benchmark, such as an index | `benchmark`, `window = 14`, `column = "close"` | `BETA` | `benchmark_<column>`, `beta_<window>` |
| `correlation_coefficient` | Adds the rolling Pearson correlation of the instrument's returns with a benchmark's returns | `benchmark`, `window = 14`, `column = "close"` | `CORREL` | `benchmark_<column>`, `corr_<window>` |
| `linear_regression` | Adds the end value of a rolling linear regression line through one candle column | `window = 14`, `column = "close"` | `LINEARREG` | `lin_regr_<window>` |
| `linear_regression_slope` | Adds the slope of a rolling linear regression line through one candle column | `window = 14`, `column = "close"` | `LINEARREG_SLOPE` | `lin_regr_slope_<window>` |
| `linear_regression_intercept` | Adds the intercept of a rolling linear regression line through one candle column | `window = 14`, `column = "close"` | `LINEARREG_INTERCEPT` | `lin_regr_int_<window>` |
| `linear_regression_angle` | Adds the angle in degrees of a rolling linear regression line through one candle column | `window = 14`, `column = "close"` | `LINEARREG_ANGLE` | `lin_regr_angle_<window>` |
| `standard_deviation` | Adds the rolling standard deviation of one candle column | `window = 14`, `standard_deviations = 1`, `column = "close"` | `STDDEV` | `std_dev_<window>` |
| `variance` | Adds the rolling variance of one candle column | `window = 14`, `standard_deviations = 1`, `column = "close"` | `VAR` | `var_<window>` |

`variance` passes `standard_deviations` to TA-Lib as it always has, but TA-Lib's variance does not use it, so the argument has no effect there.

The table below shows where each number comes from in R. The `talib` package bundles a newer TA-Lib, C 0.8.1, than the C 0.6.4 that Python's `talib` links, and the newer library rewrote `CORREL`, `VAR` and `STDDEV`, so the correlation is computed by a base R copy of the 0.6.4 loop, which also keeps 0.6.4's handling of missing values. The `talib` package has no linear regression functions at all.

| Method | Calculated by | Agreement with Python on 2026-10-07 |
|---|---|---|
| `beta` | `talib::BETA` | Relative difference of 2.3e-12 at most |
| `standard_deviation`, `variance` | `talib::STDDEV`, `talib::VAR` | Relative differences of 5.4e-12 and 1.1e-11 at most |
| `correlation_coefficient` | A base R copy of TA-Lib 0.6.4's `TA_CORREL` loop | Identical |
| the four `linear_regression` methods | A base R copy of TA-Lib 0.6.4's `TA_LINEARREG` loop, one helper for all four | Identical |

Python's `talib` skips leading rows where an input is missing and pads the result, while the `talib` package passes them to the C code, which can then give wrong or empty results. Every method here therefore starts the calculation at the first complete row and pads the front with `NA`, as Python does. This matters for `correlation_coefficient`, whose first return is always missing.

### Beta and correlation against a benchmark

`beta` and `correlation_coefficient` compare the instrument with a second one, the benchmark, which is their first and required argument. The benchmark is usually an index, such as an `EquityIndex` for NIFTY, but any object with a `prices()` method works, such as an asset basket. The sequence below shows how the two sets of candles are matched.

<pre class="mermaid">
sequenceDiagram
    autonumber
    participant Y as Your code
    participant S as share$beta
    participant U as UBI
    participant T as talib package
    Y-&gt;&gt;S: benchmark = nifty, window = 60, days = 365
    S-&gt;&gt;U: the share's candles
    S-&gt;&gt;U: the benchmark's candles, same range
    U--&gt;&gt;S: two data frames
    S-&gt;&gt;S: match on datetime, keep only candles both have
    S-&gt;&gt;T: talib::BETA(benchmark close, share close, timePeriod = 60)
    T--&gt;&gt;S: one beta per candle
    S--&gt;&gt;Y: candles with benchmark_close and beta_60
</pre>

A candle that only one side has, such as a day the stock was suspended while the index traded, is left out rather than filled in. The match keeps the instrument's row order, as pandas' inner merge does in the Python library; a benchmark with two candles at the same moment would give pandas two rows and R one. The benchmark's column is kept as `benchmark_<column>`, so you can see what was compared.

Two details make the numbers right. TA-Lib's `BETA(x, y)` returns the beta of `y` measured against `x`, which is the reverse of what its documentation suggests, so the method passes the benchmark first; the `talib` package wraps the same C function, so the same holds in R. A check of the Python library on 2026-09-14 found 2.171 for INFY against NIFTY this way, matching a hand calculation, against 0.167 the other way round. And `correlation_coefficient` correlates returns rather than price levels, because two unrelated prices that both rise would otherwise look highly correlated.

The example below, which was not run for this article, measures RELIANCE's 60-day rolling beta against NIFTY.

```{r}
reliance <- Equity$new(exchange = "nse", symbol = "RELIANCE")
nifty <- EquityIndex$new(exchange = "nse", symbol = "NIFTY")

frame <- reliance$beta(benchmark = nifty, window = 60, days = 365)
columns <- c(
  "datetime",
  "close",
  "benchmark_close",
  "beta_60"
)
print(tail(frame[, columns]))
```

## Math operators

`MathOperators` holds TA-Lib's ten math operators: arithmetic between two candle columns, and the highest and lowest values of one column over a rolling window. The table below lists them.

| Method | What it adds | Own arguments and defaults | TA-Lib function | Columns added |
|---|---|---|---|---|
| `add` | Adds the sum of two candle columns | `first_column = "high"`, `second_column = "low"` | `ADD` | `sum` |
| `subtract` | Adds the second candle column subtracted from the first | `first_column = "high"`, `second_column = "low"` | `SUB` | `difference` |
| `multiply` | Adds the product of two candle columns | `first_column = "high"`, `second_column = "low"` | `MULT` | `product` |
| `divide` | Adds the first candle column divided by the second | `first_column = "high"`, `second_column = "low"` | `DIV` | `quotient` |
| `maximum` | Adds the highest value of one candle column over each window | `column = "close"`, `window = 10` | `MAX` | `max` |
| `minimum` | Adds the lowest value of one candle column over each window | `column = "close"`, `window = 10` | `MIN` | `min` |
| `maximum_index` | Adds the row position of the highest value of one candle column over each window | `column = "close"`, `window = 10` | `MAXINDEX` | `maxindex` |
| `minimum_index` | Adds the row position of the lowest value of one candle column over each window | `column = "close"`, `window = 10` | `MININDEX` | `minindex` |
| `minimum_maximum` | Adds the lowest and highest values of one candle column over each window | `column = "close"`, `window = 10` | `MINMAX` | `min`, `max` |
| `minimum_maximum_index` | Adds the row positions of the lowest and highest values of one candle column over each window | `column = "close"`, `window = 10` | `MINMAXINDEX` | `minindex`, `maxindex` |

Three details are worth knowing. The two-column methods default to `high` and `low`, so `subtract()` with no arguments gives each candle's range. The rolling methods take `column` before `window`, the reverse of most analysis methods, so call them by name. And the index methods give positions counted from 0 for the first candle of the fetched range, exactly as the Python library does, so the row in an R data frame is the position plus 1; the rows before the first full window hold 0, because Python's `talib` fills them with 0. `minimum_maximum` writes the same `min` and `max` columns as `minimum` and `maximum`, which only matters if you combine their data frames by hand.

Every method in this group is computed in base R. The `talib` package has none of `ADD`, `SUB`, `MULT`, `DIV`, `MINMAX`, `MAXINDEX`, `MININDEX` or `MINMAXINDEX`, and its `MAX` and `MIN` come from the newer TA-Lib, which treats a missing value inside the data differently: for the values `1, 2, NA, 4, 5, 6, 7` with a window of 2, `talib::MAX` gave `NA, 2, 2, 4, 5, 6, 7` where Python's `talib.MAX` gave `NaN, 2, 2, NaN, 5, 6, 7`, checked on 2026-10-07. The four arithmetic methods use R's own operators, which are the same single operation as TA-Lib's C code, and the rolling methods follow TA-Lib 0.6.4's loops step by step, including how they break ties and treat missing values. Every value matched the Python library exactly. A window below 2 for the rolling methods signals an error naming TA-Lib's `TA_BAD_PARAM`, as Python's `talib` does.

## Math transforms

`MathTransforms` holds TA-Lib's fifteen mathematical transforms of a single column. The table below lists them.

| Method | What it adds | Own arguments and defaults | TA-Lib function | Columns added |
|---|---|---|---|---|
| `arc_cosine` | Adds the arc cosine of one candle column | `column = "close"` | `ACOS` | `acos` |
| `arc_sine` | Adds the arc sine of one candle column | `column = "close"` | `ASIN` | `asin` |
| `arc_tangent` | Adds the arc tangent of one candle column | `column = "close"` | `ATAN` | `atan` |
| `ceiling` | Adds the ceiling of one candle column | `column = "close"` | `CEIL` | `ceil` |
| `cosine` | Adds the cosine of one candle column | `column = "close"` | `COS` | `cos` |
| `hyperbolic_cosine` | Adds the hyperbolic cosine of one candle column | `column = "close"` | `COSH` | `cosh` |
| `exponential` | Adds the exponential of one candle column | `column = "close"` | `EXP` | `exp` |
| `floor` | Adds the floor of one candle column | `column = "close"` | `FLOOR` | `floor` |
| `natural_logarithm` | Adds the natural logarithm of one candle column | `column = "close"` | `LN` | `ln` |
| `logarithm_base_10` | Adds the base 10 logarithm of one candle column | `column = "close"` | `LOG10` | `log10` |
| `sine` | Adds the sine of one candle column | `column = "close"` | `SIN` | `sin` |
| `hyperbolic_sine` | Adds the hyperbolic sine of one candle column | `column = "close"` | `SINH` | `sinh` |
| `square_root` | Adds the square root of one candle column | `column = "close"` | `SQRT` | `sqrt` |
| `tangent` | Adds the tangent of one candle column | `column = "close"` | `TAN` | `tan` |
| `hyperbolic_tangent` | Adds the hyperbolic tangent of one candle column | `column = "close"` | `TANH` | `tanh` |

The `talib` package has none of TA-Lib's math transforms, so every method in this group calls the base R function that wraps the same C library function, such as `acos()` or `log10()`, which is all TA-Lib's own C code does. On the fixture candles every value was identical to Python's, including the infinities. Base R warns "NaNs produced" when `acos()`, `asin()`, `log()`, `log10()` or `sqrt()` is given a value outside its domain, where C and Python stay silent, so those five methods compute only the values inside the domain and fill the rest with `NaN`, which gives the same column without a warning.

::: {.alert .alert-warning}
**Most of these are meaningless on raw prices.**

Arc cosine and arc sine are defined only between -1 and 1, so on a price of about a thousand rupees their columns are entirely empty. `exponential`, `hyperbolic_cosine` and `hyperbolic_sine` overflow to infinity for any value above about 710: over ten years of INFY's adjusted closes, 1,586 of 1,654 rows were infinite. The transforms are meant for columns that are already small. Note that `column` must name a column of the candles themselves, so a transform of returns needs the returns added to the data frame first.
:::

## Price transforms

`PriceTransforms` holds TA-Lib's four price transforms, each of which combines a candle's prices into one representative figure. They take no arguments of their own. The table below lists them.

| Method | What it adds | Own arguments and defaults | TA-Lib function | Columns added |
|---|---|---|---|---|
| `average_price` | Adds the average of each candle's open, high, low and close | none | `AVGPRICE` | `avg_price` |
| `median_price` | Adds the midpoint of each candle's high and low | none | `MEDPRICE` | `med_price` |
| `typical_price` | Adds the average of each candle's high, low and close | none | `TYPPRICE` | `typ_price` |
| `weighted_close` | Adds each candle's weighted close, which counts the close twice alongside the high and low | none | `WCLPRICE` | `wght_close` |

`average_price` here is TA-Lib's average of each candle's open, high, low and close. It is unrelated to the `average_price` field in UBI's quote, which is the day's volume weighted average price and is read through the `volume_weighted_average_price` active binding of `TradeableInstrument`. On 2026-09-26 the Python library's property returned 1220.44 for RELIANCE.

The four price transforms call the `talib` package's `AVGPRICE`, `MEDPRICE`, `TYPPRICE` and `WCLPRICE`, and on 2026-10-07 their values matched the Python library's within a relative difference of about 1e-13.
