---
title: "Performance measures"
---

```{r, include = FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

The performance measures are the numbers used to judge an investment over a period: how much it returned, how bumpy the ride was, how far it fell, and how it did against a benchmark. They are methods on every instrument and on every [asset basket](guide-asset-baskets.html), because both supply `prices()`. They live in the [`PerformanceMeasures`](../reference/PerformanceMeasures.html) class in `R/assets_analysis_performance_measures.R`, the last link of the analysis chain, which `Instrument` and `AssetBasket` both inherit.

Each method takes the usual range arguments, `interval`, `from_date`, `to_date`, `days` and `adjusted`, and works on the closing prices. A return is the fractional change of the close from one candle to the next. Annual figures scale by 252 candles a year for `day` candles, and by the number of candles in 252 sessions of 375 minutes for a minute interval such as `5minute`; any other interval signals `ValueError` before any request is sent. A `risk_free_rate` is an annual fraction, so a 6.5 percent treasury bill is `0.065`, and it defaults to 0, because a built-in rate would go stale.

The table below lists the measures.

| Method | What it means | A good value |
|---|---|---|
| `cumulative_return` | Total growth over the range | Higher |
| `annualised_return` | The compound annual growth rate | Higher |
| `annualised_volatility` | The standard deviation of returns, scaled to a year | Lower, for the same return |
| `sharpe_ratio(risk_free_rate)` | Annual return above the risk-free rate for each unit of volatility | Above 1 is usually thought good |
| `sortino_ratio(risk_free_rate)` | The Sharpe ratio with only the falls counted as risk | Higher |
| `drawdowns` | How far below its earlier peak the close stood at every candle | A data frame, not a single number |
| `maximum_drawdown` | The worst fall from a peak, as a negative fraction | Closer to zero |
| `calmar_ratio` | Annual growth rate divided by the size of the worst drawdown | Higher |
| `value_at_risk(confidence, method)` | The one-candle loss not exceeded on that share of candles, `"historical"` or `"parametric"` | Lower |
| `expected_shortfall(confidence)` | The average loss on the candles beyond the value at risk | Lower |
| `benchmark_beta(benchmark)` | How much the price moved for each percent the benchmark moved | 1 moves with the market |
| `alpha(benchmark, risk_free_rate)` | Annual return beyond what the beta explains, Jensen's alpha | Positive |
| `tracking_error(benchmark)` | Annual volatility of the return difference from the benchmark | Near zero for an index fund |
| `information_ratio(benchmark)` | Annual return above the benchmark for each unit of tracking error | Higher |
| `up_capture_ratio(benchmark)`, `down_capture_ratio(benchmark)` | How much of the benchmark's rises and falls were captured | Up above 1, down below 1 |
| `performance_summary(benchmark = NULL, ...)` | Every measure above from one fetch of the candles | A named list |

A benchmark is anything with a `prices()` method, so an instrument can be measured against an index, a basket against the index it copies, or one fund's holdings against another's.

The R methods follow the Python library's pandas calls exactly. Standard deviations, variances and covariances divide by n - 1, as pandas does; quantiles use linear interpolation, `stats::quantile(type = 7)`; the parametric value at risk uses `stats::qnorm()`; and the running peak behind `drawdowns()` skips missing values, as pandas' `cummax()` does, where base R's `cummax()` would turn everything after a missing value into `NA`. On 2026-10-07 every measure was compared with the Python class on the same fixture candles, with risk-free rates of 0 and 0.065, `day` and `5minute` intervals, three confidence levels and both value at risk methods, and the largest relative difference was 3.8e-15.

## Example

The example below measures INFY against NIFTY over a year.

```{r}
infosys <- Equity$new(exchange = "nse", symbol = "INFY")
nifty <- EquityIndex$new(exchange = "nse", symbol = "NIFTY")
summary <- infosys$performance_summary(
  benchmark = nifty,
  risk_free_rate = 0.065,
  days = 365
)
str(summary)
```

`performance_summary()` returns a named list, where the Python library returns a pandas Series. Its names are always the same nine, or fifteen with a benchmark, in the order of the table above: `cumulative_return`, `annualised_return`, `annualised_volatility`, `sharpe_ratio`, `sortino_ratio`, `maximum_drawdown`, `calmar_ratio`, `value_at_risk` and `expected_shortfall`, followed by `benchmark_beta`, `alpha`, `tracking_error`, `information_ratio`, `up_capture_ratio` and `down_capture_ratio`. A measure that cannot be calculated is a `NULL` element that stays in the list, so `summary[["sharpe_ratio"]]` always works. It fetches the candles once for the instrument and once for the benchmark.

When the Python library ran the same summary on 2026-09-28, INFY's Sharpe ratio over the year was -1.302082 and its maximum drawdown was -0.416913, and both matched a separate calculation with pandas to every printed digit. Its beta against NIFTY was 0.684382, with an up capture ratio of 0.596877 and a down capture ratio of 0.963832.

A mutual fund returns `NULL` for every measure, because UBI stores no candles or net asset value for one; measure its [constituents](guide-asset-baskets.html#an-index-or-a-fund-is-two-things) instead.
