---
title: "Equities"
---

```{r, include = FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

The equity family covers shares, stock market indices, and the futures and options written on each of them. It lives in `R/assets_equities.R` and has six classes, one for each of UBI's equity segments. It is the family where the most works: every class that resolves an instrument has a live quote, shares and indices have candles, five of the six classes can be ordered, and `Equity` can be held.

The table below lists the six classes and what identifies one contract of each.

| Kind | Class | What it is | UBI segment | Named by |
|---|---|---|---|---|
| class | [`Equity`](#a-share) | A share, such as RELIANCE | `equities` | `exchange`, `symbol` |
| class | [`EquityFutures`](#futures-and-options) | A future on a share | `equity_futures` | `exchange`, `underlying_symbol`, `expiry_date` |
| class | [`EquityOption`](#futures-and-options) | An option on a share | `equity_options` | `exchange`, `underlying_symbol`, `expiry_date`, `strike_price`, `option_type` |
| class | [`EquityIndex`](#an-index) | A stock market index, such as NIFTY | `equity_indices` | `exchange`, `symbol` |
| class | [`EquityIndexFutures`](#futures-and-options) | A future on an index | `equity_index_futures` | `exchange`, `underlying_symbol`, `expiry_date` |
| class | [`EquityIndexOption`](#futures-and-options) | An option on an index | `equity_index_options` | `exchange`, `underlying_symbol`, `expiry_date`, `strike_price`, `option_type` |

## How the six classes are built

`EquityIndex` is built on [`NonTradeableInstrument`](../reference/NonTradeableInstrument.html), because an index is a number the exchange publishes rather than something you can buy. The other five can be traded, so they carry the order book, the order methods, the price wrappers and the position members. `Equity` is built on [`TradeableInstrument`](../reference/TradeableInstrument.html) directly and alone adds the holdings members. The four derivative classes are built on the [derivative base classes](guide-derivatives.html), which add the expiry, underlying, basis and greeks members, so each derivative class adds only its segment and its discovery functions. The discovery functions live on each class's generator, such as `EquityOption$chain()`, because R6 generators do not inherit functions from their parents. The class diagram below shows this.

<pre class="mermaid">
classDiagram
    direction BT
    class Instrument {
        prices()
        quote
        last_price
        ohlc
        209 analysis methods
    }
    class TradeableInstrument {
        order book values
        place_order()
        price wrappers
        position members
    }
    class Equity {
        holdings members
        generator search()
    }
    class NonTradeableInstrument {
        constituents
    }
    class EquityIndex {
        generator search()
    }
    class Derivative {
        expiry and underlying members
    }
    class Futures {
        basis members
    }
    class Option {
        moneyness, greeks()
    }
    class IndexFutures
    class IndexOption
    class EquityFutures {
        generator expiries()
        generator contracts()
    }
    class EquityIndexFutures {
        generator expiries()
        generator contracts()
    }
    class EquityOption {
        generator expiries()
        generator strikes()
        generator chain()
    }
    class EquityIndexOption {
        generator expiries()
        generator strikes()
        generator chain()
    }
    TradeableInstrument --|&gt; Instrument
    NonTradeableInstrument --|&gt; Instrument
    Equity --|&gt; TradeableInstrument
    Derivative --|&gt; TradeableInstrument
    Futures --|&gt; Derivative
    Option --|&gt; Derivative
    IndexFutures --|&gt; Futures
    IndexOption --|&gt; Option
    EquityFutures --|&gt; Futures
    EquityOption --|&gt; Option
    EquityIndexFutures --|&gt; IndexFutures
    EquityIndexOption --|&gt; IndexOption
    EquityIndex --|&gt; NonTradeableInstrument
</pre>

## Naming a contract

Every identity argument is required, and the constructor accepts only the fields that identify one of its own contracts. The four derivative classes also take an optional `underlying`, an instrument object the contract keeps, as [How a derivative finds its underlying](#how-a-derivative-finds-its-underlying) explains. Every constructor also takes an optional `unified_broker_interface`, the client to send requests through, which defaults to the one client all instruments share. A share takes an exchange and a symbol, a future adds nothing to that but swaps the symbol for its underlying's symbol and an expiry date, and an option also takes a strike price and an option type, `CE` for a call or `PE` for a put. The exchange is written in lower case, such as `nse` or `bse`.

Leaving out a field fails at once, in R, before any request is sent: R signals its own error, such as `argument "symbol" is missing, with no default`. Asking for something UBI does not have fails one request later with the class's own error condition, such as `EquityOptionError`, whose `parent` field holds the general `InstrumentError` carrying UBI's own message. Asking one class for another's instrument, such as `Equity$new(exchange = "nse", symbol = "NIFTY")`, fails the same way, because the class fixes its segment and UBI finds no share called NIFTY.

The code below builds one contract of each kind. It is the module's own usage example from the Python library, translated to R and extended to the futures classes.

```{r}
library(tradeR)

share <- Equity$new(exchange = "nse", symbol = "RELIANCE")
nifty <- EquityIndex$new(exchange = "nse", symbol = "NIFTY")

future <- EquityFutures$new(
  exchange = "nse",
  underlying_symbol = "RELIANCE",
  expiry_date = "2026-09-29"
)
option <- EquityIndexOption$new(
  exchange = "nse",
  underlying_symbol = "NIFTY",
  expiry_date = "2026-09-29",
  strike_price = 25000,
  option_type = "CE"
)
```

## A share

An [`Equity`](../reference/Equity.html) is a share on the nse or the bse. The example below builds RELIANCE and reads its identity fields, its last ten days of candles and its last price.

```{r}
reliance <- Equity$new("nse", "RELIANCE")
print(reliance)
reliance$instrument_id
reliance$segment
reliance$lot_size
reliance$tick_size
reliance$carried_by
reliance$prices(days = 10)
reliance$last_price
```

The output of the Python version of this example was captured from a local UBI on Saturday 2026-09-26, when the market was closed, so the price was Friday's close. It is not shown here, because R prints differently, but the values it recorded are described below. The instrument printed as `Equity(exchange='nse', segment='nse_equities', symbol='RELIANCE')`. Its `instrument_id` was `3f92570a-9924-5bf5-9f9d-e006cd9f4202`, its `shape` was `security`, its `lot_size` was 1 and its `tick_size` was 0.1, and its `carried_by` list, which names every broker that carries the share, had nine entries, such as Dhan with broker token `2885` and Zerodha with broker token `738561` and order symbol `RELIANCE`. The capture's `mapping_date` was 2026-09-26, its `first_seen_date` 2026-08-07 and its `last_seen_date` 2026-09-26. The candles came back as six daily rows starting on 2026-09-16, with the columns `exchange`, `segment`, `interval`, `datetime`, `open`, `high`, `low`, `close`, `volume`, `oi` and `price_factor`, and the last price was 1226.0.

The candles carry a `price_factor` column because a share's prices are adjusted for splits and bonuses by default; [`prices()`](guide-market-data.html#prices) explains the columns and UBI's page on [adjusted and unadjusted prices](https://pramodathani.github.io/unified_broker_interface/rest-api/historical-data/#adjusted-unadjusted-and-as-served) explains the adjustment. The same capture asked for `prices(interval = "5minute", days = 1)` and got nothing, which is `NULL` in R, because intraday candles are loaded into UBI by hand and none had been loaded for RELIANCE.

### Holding a share

`Equity` is the only class in this family with the holdings members, because a share is the only equity contract that can be kept in the demat account. A future or an option leaves a position rather than a holding, and an index cannot be held at all. The table below lists the six members, which are documented in full on the [Holdings](guide-holdings.html) page.

| Kind | Member | Description |
|---|---|---|
| active binding | `holdings` | This share's row from the account's holdings, or `NULL` when it is not held |
| active binding | `holdings_value` | What the holding is worth, as UBI prices it |
| active binding | `holdings_pnl` | The holding's `day_change`, `day_change_percentage` and `unrealized` profit |
| places orders | `add_to_holdings()` | Buys more, always as a `cnc` order |
| places orders | `reduce_holdings()` | Sells part of the shares that are free to sell |
| places orders | `liquidate_holdings()` | Sells every share that is free to sell |

Two details matter in practice. The three order methods always send the `cnc` product, because selling a holding as `mis` would open an intraday short position beside the shares instead of selling them. And the shares free to sell are the holding minus any shares pledged as collateral, so `reduce_holdings()` refuses a quantity larger than that with `HoldingError`. The RELIANCE capture above found RELIANCE not held, so `holdings` returned nothing, which is `NULL` in R.

## An index

An [`EquityIndex`](../reference/EquityIndex.html) has a live level and candles, so every analysis method works on it, but it has no order book and no order methods. It is also the usual benchmark for [`beta()`](analysis-statistics.html) and for the [performance measures](analysis-performance.html), and the Python library checked `beta` against NIFTY's candles on 2026-09-14. Its [`constituents`](guide-asset-baskets.html#an-index-or-a-fund-is-two-things) active binding, which every index class inherits from `NonTradeableInstrument`, returns the stored basket of the index's members, or `NULL` when none has been stored; UBI stores no constituents, so they are kept in this project's MongoDB. The example below searches the nse's indices for names containing NIFTY, with `limit = 5`.

```{r}
matches <- EquityIndex$search("nse", "NIFTY", limit = 5)
print(matches)
```

When the Python version was captured from a local UBI on 2026-09-26, the search returned five rows, each with the columns `exchange`, `expiry_date`, `instrument_id`, `option_type`, `segment`, `shape`, `strike_price`, `symbol` and `underlying_symbol`. The five symbols were `NIFTY`, `NIFTY 100`, `NIFTY 200`, `NIFTY 500` and `NIFTY ALPHA 50`, all in the segment `nse_equity_indices` with the shape `security`; the first, `NIFTY`, had the `instrument_id` `dba60324-760b-53cc-aeae-a4bb4defd1bc`. In R, the result is a data frame with the same columns, and the fields that do not apply to an index are `NA`.

The index the exchange publishes as "NIFTY 50" is stored by UBI as `NIFTY`, which is the name its futures and options use as their underlying symbol. The section on [how a derivative finds its underlying](#how-a-derivative-finds-its-underlying) explains why that matters.

The same search on shares, `Equity$search("nse", "RELI", limit = 5)`, returned four rows in the same capture: RELIABLE, RELIANCE, RELIGARE and RELINFRA.

## Futures and options

The four derivative classes are named by their underlying symbol, their expiry date and, for options, a strike and an option type. You rarely know those by heart, so each class's generator offers discovery functions that read UBI's full instrument list for its own segment and return what is live. [Finding instruments](guide-discovery.html) documents them; the table below says which class has which.

| Class | `expiries` | `contracts` | `strikes` | `chain` |
|---|:-:|:-:|:-:|:-:|
| `EquityFutures`, `EquityIndexFutures` | yes | yes | no | no |
| `EquityOption`, `EquityIndexOption` | yes | no | yes | yes |

The example below walks from the NIFTY index to its option chain for the nearest expiry, and then lists the expiries of the RELIANCE futures.

```{r}
expiries <- EquityIndexOption$expiries("nse", "NIFTY")
strikes <- EquityIndexOption$strikes("nse", "NIFTY", expiries[1])
chain <- EquityIndexOption$chain("nse", "NIFTY", expiries[1])

print(EquityFutures$expiries("nse", "RELIANCE"))
```

When the Python version was captured from a local UBI on 2026-09-26, the NIFTY options were listed for eighteen expiries, from 2026-09-29 through weekly and monthly dates to 2031-06-24. The nearest expiry had 269 strikes, from 1500 to 49500, and its chain had 538 rows, which are the 269 strikes as calls and puts, sorted by strike and then option type, so the first rows were the 1500 call and the 1500 put. The RELIANCE futures were listed for three expiries: 2026-09-29, 2026-10-27 and 2026-11-23. In R, `expiries` is a `Date` vector, `strikes` is a numeric vector and `chain` is a data frame.

The discovery functions return rows of identities, not instrument objects. Building 538 objects would send 538 lookups to UBI, so you take the two or three rows you want and build those. A row's identity fields rebuild the same instrument: on 2026-09-20 a row from the middle of the RELIANCE chain was turned into an `EquityOption`, and UBI returned the very `instrument_id` the row carried.

Single-stock options are slower to discover than index options. Each call downloads the whole `nse_equity_options` segment, which held 125,967 rows on 2026-09-20 and took about two seconds, against a quarter of a second for the 14,826 rows of index options.

The six contracts below were built and checked against UBI on 2026-09-20, which shows the lot and tick sizes you can expect. The prices are that day's and are here only to show that each class was quoted.

| Class | Contract | Segment | `lot_size` | `tick_size` | `last_price` |
|---|---|---|---|---|---|
| `Equity` | RELIANCE | `nse_equities` | 1 | 0.1 | 1226.4 |
| `EquityFutures` | RELIANCE 2026-09-29 | `nse_equity_futures` | 500 | 0.1 | 1242.0 |
| `EquityOption` | RELIANCE 2026-09-29 1250 CE | `nse_equity_options` | 500 | 0.05 | 11.95 |
| `EquityIndex` | NIFTY | `nse_equity_indices` | 1 | 0.05 | 23346.4 |
| `EquityIndexFutures` | NIFTY 2026-09-29 | `nse_equity_index_futures` | 65 | 0.1 | 23380.0 |
| `EquityIndexOption` | NIFTY 2026-09-29 23350 CE | `nse_equity_index_options` | 65 | 0.05 | 171.1 |

An equity order is a securities-market order, so its quantity is a plain count of shares, and for a derivative it must be a whole number of the chosen broker's lot size. [Orders](guide-orders.html#place_order) explains the rest.

Once built, a contract reports what it is as a contract: its days to expiry, whether it is a weekly or monthly expiry, its underlying's price, a future's basis over it, and an option's moneyness, implied volatility and greeks. Equities are the one family where all of these work, because UBI quotes shares and indices; [Derivatives](guide-derivatives.html) documents each member with values captured from NIFTY and RELIANCE contracts.

## How a derivative finds its underlying

A RELIANCE option does not hold its RELIANCE `Equity` unless you give it one. Give it when you build the contract, as `EquityOption$new("nse", "RELIANCE", "2026-10-27", 1200, "CE", underlying = reliance)`, and its [`underlying`](guide-derivatives.html#underlying) returns that very object and its [`underlying_price`](guide-derivatives.html#underlying_price) reads that object's last price. Only a given object is stored. Without one, the contract looks its underlying up again on every read, in the order [Derivatives](guide-derivatives.html#how-a-contract-finds-its-underlying) describes. It first tries UBI's `underlying_instrument_id`, which UBI resolves from the brokers' own underlying codes, and then falls back to the equity family's default, which is the share or index whose `symbol` equals the contract's `underlying_symbol`. Either way the lookup returns a plain `TradeableInstrument` or, for an index, a `NonTradeableInstrument`, never an `Equity` or an `EquityIndex`. Giving the underlying is still the reliable way, because it costs no request and cannot fail.

The fallback matches strings, not keys. The flowchart below shows how that string match works for a share and for an index, and where it can fail.

<pre class="mermaid">
flowchart LR
    O["EquityOption&lt;br/&gt;underlying_symbol RELIANCE"] --&gt;|"same string"| S["Equity&lt;br/&gt;symbol RELIANCE"]
    I["EquityIndexOption&lt;br/&gt;underlying_symbol NIFTY"] --&gt;|"same string"| N["EquityIndex&lt;br/&gt;symbol NIFTY"]
    P["Exchange name&lt;br/&gt;NIFTY 50"] --&gt;|"UBI alias table"| N
    Q["An index missing&lt;br/&gt;from the alias table"] -.-&gt;|"may not match"| X["no EquityIndex found"]
</pre>

For shares the match holds, because UBI strips the exchange's series suffix, such as `-EQ`, from NSE symbols when it builds its instrument list. For indices it depends on an alias table in UBI that rewrites the published names onto the derivative names, so Zerodha's `NIFTY 50` row is stored as `NIFTY` and its `NIFTYBANK` row as `BANKNIFTY`. The table covers NIFTY, BANKNIFTY, FINNIFTY, MIDCPNIFTY and NIFTYNXT50. An index outside it may not match by name, and when UBI gives no link either, the lookup signals [`UnderlyingError`](guide-errors.html#underlyingerror) at the moment it is read, which is why passing the `EquityIndex` yourself is safer. The live check of the Python library on 2026-09-28 resolved both RELIANCE and NIFTY either way.

## Errors

The table below lists what each constructor signals. Every error condition except R's own missing-argument error carries the class `InstrumentError` as well as its own, so `tryCatch(..., InstrumentError = function(error) ...)` catches them all, and [Errors](guide-errors.html) documents each class.

| Condition | When |
|---|---|
| R's error `argument "..." is missing, with no default` | A required identity argument is missing. R signals it before any request is sent. |
| `EquityError` | UBI has no share with that exchange and symbol, including an index asked for as a share. |
| `EquityFuturesError` | UBI has no share future with that underlying and expiry. |
| `EquityOptionError` | UBI has no share option with those five fields, such as an impossible strike. |
| `EquityIndexError` | UBI has no index with that exchange and symbol, including a share asked for as an index. |
| `EquityIndexFuturesError` | UBI has no index future with that underlying and expiry. |
| `EquityIndexOptionError` | UBI has no index option with those five fields. |
