---
title: "Installation"
---

```{r, include = FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

This page installs the package and the services around it, in the order they depend on each other. The system libraries come first because several R packages fail to compile without them; the R packages the package imports come next; the package itself after that; then the MongoDB container; and last comes the UBI that every call goes to.

The table below summarises the steps and what each one leaves behind.

| Step | What you run | What you end up with |
|---|---|---|
| [1. The system libraries](#install-the-system-libraries) | `sudo apt install ...` | The C libraries and CMake that `httr2`, `mongolite` and `talib` compile against |
| [2. The R packages](#install-the-r-packages) | `install.packages(...)` | The six packages tradeR imports, plus the development tools |
| [3. The package](#install-the-package) | `devtools::install()` | `library(tradeR)` working from any R session |
| [4. MongoDB](#start-the-containers) | `docker compose up -d` in the Python project | Redis, MongoDB and TimescaleDB on ports 2002 to 2004 |
| [5. UBI](#make-sure-ubi-and-its-order-engine-are-running) | UBI's own services | UBI answering on `127.0.0.1:8080`, placing orders through its engine |

The diagram below shows what is running on the machine once every step is done, and which of those pieces the package talks to.

<pre class="mermaid">
flowchart LR
    subgraph rlib["The R library"]
        LIB["tradeR&lt;br/&gt;installed package"]
        TA["talib package"] --&gt; C["bundled TA-Lib C library"]
        LIB --&gt; TA
    end
    subgraph compose["docker compose, in tradingmachine"]
        R[("Redis&lt;br/&gt;port 2002")]
        M[("MongoDB&lt;br/&gt;port 2003")]
        T[("TimescaleDB&lt;br/&gt;port 2004")]
    end
    subgraph ubi["UBI, a separate project"]
        API["REST API&lt;br/&gt;127.0.0.1:8080"] --&gt; ENG["order engine"]
    end
    LIB -- "UBI's key and secret,&lt;br/&gt;stored baskets" --&gt; M
    LIB -- "every price and order" --&gt; API
</pre>

The `.env` file that steps 4 and 5 read is described on [Configuration](get-started-configuration.html). Write it before step 4 if you are setting up from nothing.

## 1. Install the system libraries

Several of the R packages tradeR needs are compiled from source on Linux, and they compile against C libraries that R cannot install itself. Without them the install fails with an error about a missing header such as `curl/curl.h` or `sasl/sasl.h` rather than about a missing package, which is confusing the first time.

The `talib` package, which the analysis methods use, is the R binding of the TA-Lib C library. Unlike the Python wrapper, it bundles the C library and builds it with CMake, so there is no separate TA-Lib install, but CMake must be present.

On Ubuntu the system libraries and CMake are installed with the command below, which is the one in the package's README.

```bash
sudo apt install -y libcurl4-openssl-dev libssl-dev libsasl2-dev libuv1-dev libgit2-dev libicu-dev libharfbuzz-dev libfribidi-dev libfreetype-dev libpng-dev libtiff-dev libjpeg-dev libfontconfig1-dev libxml2-dev cmake
```

## 2. Install the R packages

The package imports six R packages, and the development tools are three more. The command below installs all of them from CRAN in one call, run from any R session.

```{r}
install.packages(
  c(
    "R6",
    "httr2",
    "jsonlite",
    "mongolite",
    "dotenv",
    "talib",
    "devtools",
    "testthat",
    "withr"
  )
)
```

The table below lists what that brings in. `DESCRIPTION` declares the first six under `Imports`, because the package's code calls them, and `testthat` and `withr` under `Suggests`, because only the tests use them; `devtools` is not declared at all, because only the developer's own commands use it.

| Group | Packages | What they do here |
|---|---|---|
| The package itself, `Imports` | `R6`, `httr2`, `jsonlite`, `mongolite`, `dotenv`, `talib` | The classes, the HTTP client, JSON, MongoDB, the `.env` file and the TA-Lib indicators |
| The tests, `Suggests` | `testthat`, `withr` | Running the test suite with fake clients and mocked responses |
| The development tools | `devtools` | Installing from source, loading the package while working on it, and running the tests |

The tidyverse is deliberately not needed, because the package is written in base R.

## 3. Install the package

The package is installed from a copy of its repository. The command below, run in R from the repository root, builds the package and installs it into your R library, so that `library(tradeR)` works from any R session afterwards.

```{r}
devtools::install()
```

A quick check that the package and its TA-Lib binding are both in place is shown below. Loading the package touches nothing outside the R process, because its configuration is read only when first used.

```{r}
packageVersion("tradeR")
packageVersion("talib")
```

On the development machine on 2026-10-07 these printed `0.1.0` and `0.9.4`, under R 4.6.1.

::: {.alert .alert-info}
**Working on the package without reinstalling it.**

When you change the source in `R/`, `devtools::load_all()` loads the package straight from the repository, so an edit takes effect without running `devtools::install()` again. The test suite, run with `devtools::test()`, drives every class through a fake client and mocked HTTP responses, so it never reaches UBI and never places an order.
:::

## 4. Start the containers

The package needs a MongoDB, and it uses the same one as the Python library `tradingmachine`, because both read the same `.env` file. This repository has no Docker Compose file of its own: the containers are defined in the `docker-compose.yml` of the Python project, which runs three data stores, each in its own container, and reads their ports and passwords from that project's `.env` file. If the Python library is already set up on the machine, its containers are already the ones this package uses, and this step is done.

The commands below start the containers and list them, run from the Python project's directory.

```bash
cd ~/Projects/tradingmachine
docker compose up -d
docker compose ps
```

The table below lists the three services. Only MongoDB is used by this package, for UBI's key and secret and for the stored [asset baskets](guide-asset-baskets.html); Redis and TimescaleDB are running for code that has not been written yet.

| Service | Image | Host port | Container port | Volume | Used by the package |
|---|---|---|---|---|---|
| `redis` | `redis:trixie` | 2002 | 6379 | `tradingmachine_redis_volume` | No |
| `mongodb` | `mongo:8.0.4` | 2003 | 27017 | `tradingmachine_mongodb_volume` | Yes |
| `timescaledb` | `timescale/timescaledb:latest-pg18` | 2004 | 5432 | `tradingmachine_timescaledb_volume` | No |

The output below is `docker compose ps` on the development machine on 2026-09-26, taken from the Python project's documentation. The three warnings about `PYTHONPATH` are harmless: a leftover line in that project's `.env` refers to a variable that is not set, and no container uses it.

```text
time="2026-09-26T09:55:16+05:30" level=warning msg="The \"PYTHONPATH\" variable is not set. Defaulting to a blank string."
time="2026-09-26T09:55:16+05:30" level=warning msg="The \"PYTHONPATH\" variable is not set. Defaulting to a blank string."
time="2026-09-26T09:55:16+05:30" level=warning msg="The \"PYTHONPATH\" variable is not set. Defaulting to a blank string."
NAME                           IMAGE                               COMMAND                  SERVICE       CREATED       STATUS                 PORTS
tradingmachine-mongodb-1       mongo:8.0.4                         "docker-entrypoint.s…"   mongodb       11 days ago   Up 8 hours (healthy)   0.0.0.0:2003->27017/tcp
tradingmachine-redis-1         redis:trixie                        "docker-entrypoint.s…"   redis         11 days ago   Up 8 hours (healthy)   0.0.0.0:2002->6379/tcp
tradingmachine-timescaledb-1   timescale/timescaledb:latest-pg18   "docker-entrypoint.s…"   timescaledb   11 days ago   Up 8 hours (healthy)   0.0.0.0:2004->5432/tcp
```

Ports 2002 to 2004 were chosen so the project can run beside UBI, whose own containers use ports 1002 to 1005 on the same machine. The ports are bound on `0.0.0.0`, so other machines on the local network can reach them too.

::: {.alert .alert-danger}
**`docker compose down -v` deletes the data.**

`docker compose down` stops the containers and keeps their volumes. Adding `-v` also deletes the three `tradingmachine_*_volume` volumes and everything in them, including the MongoDB document holding UBI's key and secret and every stored asset basket, which both this package and the Python library read.
:::

The usernames, passwords and database names in `.env` are read only when a container starts with an empty volume. Changing them later does not change the existing database; change the credential inside the database, or delete the volume and start again.

## 5. Make sure UBI and its order engine are running

UBI is a separate project with its own installation, documented on [its own site](https://pramodathani.github.io/unified_broker_interface/get-started/). This package needs two things from it, listed below.

1. **UBI's REST API answers on `http://127.0.0.1:8080`.** UBI binds to `127.0.0.1` by default, so it is reachable only from the same machine, unlike the databases.
2. **UBI's order engine service is running.** UBI hands every order to a separate process, `unified-orders@order_engine.service`, and has done so for every order since it removed its direct placement mode on 2026-09-27. While the process is stopped, UBI refuses every order with HTTP 503, which the package signals as [`ServiceUnavailableError`](guide-errors.html#serviceunavailableerror).

The output below shows the two checks on the development machine on 2026-09-26. UBI's greeting route needs no token, so `curl` can call it directly.

```bash
curl http://127.0.0.1:8080/api/
systemctl --user is-active unified-rest-api.service unified-orders@order_engine.service
```

```text
{"message":"Welcome to the Unified Broker Interface API"}
active
active
```

UBI's [Services](https://pramodathani.github.io/unified_broker_interface/operations/services/#the-order-engine-and-the-synthetic-order-book) page shows how to start the order engine, and its [Order engine](https://pramodathani.github.io/unified_broker_interface/rest-api/order-engine/) page explains what it does.

::: {.alert .alert-danger}
**The order engine places real orders.**

Once the engine runs, a synthetic order can place real orders long after the call that created it has returned. Stopping UBI's REST API does not stop the engine.
:::

## What comes next

With everything installed, [Configuration](get-started-configuration.html) writes the `.env` file and the MongoDB document the package reads, and [First steps](get-started-first-steps.html) makes the first calls.
