---
title: "Placing and reading orders"
---

```{r, include = FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

Every instrument that can be traded carries the members on this page. Five of them send orders to UBI, which passes them on to a broker, six read this instrument's own rows out of the day's order book and trade book, and four read the order engine's parents, which are the synthetic orders and held limit orders UBI is working. An index cannot be traded, so `EquityIndex` and the other index classes have none of these members. Every member on this page is defined on [TradeableInstrument](../reference/TradeableInstrument.html), and every tradeable class, such as `Equity` or `CommodityFutures`, inherits it.

::: {.alert .alert-danger}
**These are real orders.**

The members marked "places orders" send orders through UBI to real brokers, with real money. UBI does not check the market's hours, and a broker that accepts an order can still see it filled within milliseconds. Pass `dry_run = TRUE` to `place_order()`, `modify_order()` or `cancel_order()` first: UBI then builds the exact request it would send to the broker and returns it without sending anything.
:::

The table below lists the fifteen members this page documents.

| Kind | Member | Description |
|---|---|---|
| places orders | [`place_order()`](#place_order) | Places one order in this instrument through UBI. |
| places orders | [`modify_order()`](#modify_order) | Changes one pending order through UBI. |
| places orders | [`cancel_order()`](#cancel_order) | Cancels one pending order through UBI. |
| places orders | [`cancel_open_orders()`](#cancel_open_orders) | Cancels every order in this instrument that is still waiting, whether at a broker or held in UBI's order engine. |
| places orders | [`cancel_parent()`](#cancel_parent) | Cancels one of the order engine's parents, with every leg it still has resting at a broker. |
| active binding | [`orders`](#orders) | Every one of today's orders in this instrument, whatever its status. |
| active binding | [`open_orders`](#open_orders) | Today's orders in this instrument that can still be changed. |
| active binding | [`completed_orders`](#completed_orders) | Today's orders in this instrument that filled in full. |
| active binding | [`rejected_orders`](#rejected_orders) | Today's orders in this instrument that a broker or the exchange refused. |
| active binding | [`cancelled_orders`](#cancelled_orders) | Today's orders in this instrument that were cancelled. |
| active binding | [`trades`](#trades) | Today's trades in this instrument. |
| active binding | [`parents`](#parents) | This instrument's synthetic orders and held orders that the order engine has not finished. |
| method | [`parent()`](#parent) | Reads one of the order engine's parents, whether or not it has finished. |
| method | [`parent_orders()`](#parent_orders) | Today's broker orders that one parent placed. |
| method | [`parent_trades()`](#parent_trades) | Today's trades in the broker orders that one parent placed. |

Placing an order through a named price, such as "buy at the best bid", has its own page, [Price wrappers](guide-price-wrappers.html). Changing a position without saying which side you are on is on [Positions](guide-positions.html), and the fifty-four order types UBI builds out of ordinary orders are on [Synthetic orders](guide-synthetic-orders.html). All of them end in `place_order()`.

## Glossary of plain strings

The package passes the order vocabulary to UBI as plain lower-case strings, with no constants and no factor levels, because UBI already checks every value and answers HTTP 400 with a clear message when one is wrong. UBI ignores case, so `"buy"` and `"BUY"` are the same. The table below lists the values these members accept and return; [Vocabulary](guide-vocabulary.html) has the full list for the whole package.

| Parameter | Values | Meaning |
|---|---|---|
| `transaction_type` | `buy`, `sell` | The side of the order. |
| `order_type` | `market` | Fill at whatever price the market offers. Takes no price. |
| | `limit` | Fill at the price given or better. Needs a price. |
| | `sl` | A stop-loss limit: waits for the trigger price, then rests a limit. Needs a price and a trigger price. |
| | `sl-m` | A stop-loss market: waits for the trigger price, then sends a market order. Needs a trigger price and takes no price. |
| `product` | `cnc` | Delivery. A buy puts shares in the demat account. |
| | `mis` | Intraday. The broker closes the position before the session ends. |
| | `nrml` | Carry forward, for futures and options held overnight. |
| `validity` | `day`, `ioc` | Good for the day, or immediate-or-cancel. UBI uses `day` when none is given. |
| `status` (returned) | `PENDING`, `OPEN`, `COMPLETE`, `CANCELLED`, `REJECTED`, `EXPIRED` | UBI's own upper-case status of an order. |
| `outcome` (returned) | `accepted`, `rejected`, `unknown`, `partial`, `armed` | What happened to the request. `partial` means some of the orders of a type that sends several at once were accepted and some were not. `armed` means the order engine is holding the order, waiting for a price or a time. |

UBI ties the price fields to the order type, and the package does not check this before sending. The table below shows the rule, which you will otherwise meet as a `BadRequestError`.

| `order_type` | `price` | `trigger_price` |
|---|:---:|:---:|
| `market` | must be absent | must be absent |
| `limit` | required | must be absent |
| `sl` | required | required |
| `sl-m` | must be absent | required |

A `price_reference` stands in for `price`, and a `quantity_reference` stands in for `quantity`. Both are described under [`place_order()`](#place_order).

## Nothing is checked before sending

The package sends the price and the quantity to UBI exactly as you give them. It does not round a price to the tick size, it does not check that a quantity is a whole number of lots, and it does not check that the market is open. UBI and the broker behind it hold those rules, and their refusal is the most accurate message you can get. The consequence is that the error you see comes from UBI, as one of the conditions on the [Errors](guide-errors.html) page.

Quantities are always in units, never in lots. An MCX gold future has a lot of 100, so one lot is `quantity = 100`, and `quantity = 1` is refused with HTTP 400.

## How an order travels

The sequence below follows one order that carries a `price_reference`, from your code to the broker and back. Every order takes this path through UBI's order engine, which has placed every order since UBI removed its direct placement mode on 2026-09-27.

<pre class="mermaid">
sequenceDiagram
    autonumber
    participant P as Your program
    participant I as TradeableInstrument
    participant C as Shared client
    participant A as UBI API worker
    participant E as UBI order engine
    participant B as Broker
    P-&gt;&gt;I: place_order with a price_reference
    I-&gt;&gt;C: the body
    C-&gt;&gt;A: POST /api/orders/place
    A-&gt;&gt;A: check the body and resolve the instrument
    A-&gt;&gt;E: intent on the Redis stream
    E-&gt;&gt;E: read the quote, work the price out, round it to the tick
    E-&gt;&gt;B: the broker's own order request
    B--&gt;&gt;E: order id
    E--&gt;&gt;A: answer on the intent's list
    A--&gt;&gt;C: 200 with order_id, intent_id and parent_id
    C--&gt;&gt;I: answer
    I--&gt;&gt;P: the answer as a named list
</pre>

A plain `day` limit order with a price of its own stops at step 6: the engine holds it, answers at once with HTTP 202 and an `outcome` of `armed`, and sends it only when the book reaches its price. [Order engine](architecture-order-engine.html#plain-limit-orders-are-held) explains that, and UBI's page on the [order engine](https://pramodathani.github.io/unified_broker_interface/rest-api/order-engine/) describes the engine itself.

## place_order

The method's signature is shown below. It places orders, and it sends `POST /api/orders/place`.

```r
place_order(
  transaction_type,
  order_type,
  quantity,
  product,
  price = NULL,
  trigger_price = NULL,
  validity = NULL,
  disclosed_quantity = NULL,
  after_market = FALSE,
  tag = NULL,
  dry_run = FALSE,
  price_reference = NULL,
  quantity_reference = NULL,
  synthetic = NULL
)
```

This method places one order in this instrument. You never name a broker: UBI chooses one from the brokers that carry the instrument and can take the order, and the answer says which it chose. Every optional field left as `NULL` is left out of the request entirely, rather than sent as zero, because UBI treats a missing price and a zero price differently.

An answer with an `outcome` of `accepted` means the broker took the order, not that the order survived. The exchange can still refuse it a moment later, which is what happens to an ordinary order sent while the market is closed, so read its real fate from [`orders`](#orders). To queue an order for the next session, pass `after_market = TRUE`.

The last three parameters describe something for UBI to work out instead of stating it. A `price_reference` names a price, such as "the second best offer", which UBI reads from the live quote and rounds to the tick when it sends the order. A `quantity_reference` names a quantity, such as "the whole position", which UBI reads from the account's positions. A `synthetic` named list turns the order into one of UBI's fifty-four synthetic order types.

A plain `limit` order with a price of its own, `day` validity and no `synthetic` list is held by UBI's order engine rather than sent, until the other side of the book reaches its price. The answer then has an `outcome` of `armed`, a `parent_id` and no `order_id`, and the order is found in [`parents`](#parents) rather than in `orders`. Pass `synthetic = list(type = "simple")` to send a limit order at once, which an instrument with no live quote needs, because its held order would never be sent.

A plain `market` order that is not after-market is not sent as a market order either. The engine runs it as a `marketable_limit`, a `limit` two ticks past the other side's best price that follows that price until it fills and is cancelled after 30 seconds, and refuses it with HTTP 409 when nobody is on the other side of the book or the quote is missing or stale. `synthetic = list(type = "simple")` sends a real market order. [Order engine](architecture-order-engine.html#market-orders-are-sent-as-marketable-limits) explains it.

#### Parameters

The table below lists the method's parameters.

| Name | Type | Required | Default | Description |
|---|---|:---:|---|---|
| `transaction_type` | character | yes | | `buy` or `sell`. UBI overrides it for a `quantity_reference` that reduces or closes a position. |
| `order_type` | character | yes | | `market`, `limit`, `sl` or `sl-m`. |
| `quantity` | integer or `NULL` | yes | | The quantity in units, not lots, or `NULL` when a `quantity_reference` supplies it. It has no default, so you have to write `NULL` on purpose. |
| `product` | character | yes | | `cnc`, `mis` or `nrml`. |
| `price` | numeric or `NULL` | no | `NULL` | The limit price in rupees. Leave it out for `market` and `sl-m`, or when a `price_reference` supplies it. |
| `trigger_price` | numeric or `NULL` | no | `NULL` | The trigger price in rupees, for `sl` and `sl-m`. |
| `validity` | character or `NULL` | no | `NULL` | `day` or `ioc`. UBI uses `day` when it is `NULL`. |
| `disclosed_quantity` | integer or `NULL` | no | `NULL` | The part of the order to show on the exchange. `NULL` shows all of it. |
| `after_market` | logical | no | `FALSE` | `TRUE` sends an after-market order, which the broker queues for the next session. |
| `tag` | character or `NULL` | no | `NULL` | A label of up to twenty letters and digits. |
| `dry_run` | logical | no | `FALSE` | `TRUE` has UBI build the broker's request and return it without sending it. |
| `price_reference` | named list or `NULL` | no | `NULL` | A price for UBI to work out, such as `list(kind = "offer_level", level = 2)`. The seven kinds are on [Price wrappers](guide-price-wrappers.html#the-seven-price-references). |
| `quantity_reference` | named list or `NULL` | no | `NULL` | A quantity for UBI to work out, such as `list(kind = "liquidate_position", product = "intraday")`. The kinds are on [Positions](guide-positions.html#the-quantity-reference-ubi-resolves). |
| `synthetic` | named list or `NULL` | no | `NULL` | A synthetic order type and its settings, such as `list(type = "bracket", stop_price = 990, stop_limit_price = 988, target_price = 1010)`. The classes on [Synthetic orders](guide-synthetic-orders.html) build it for you. |

#### Example

The first example below is a dry run of a plain limit buy of one RELIANCE share at 1000 rupees. The second asks UBI to price the order at the best offer instead of stating a price.

```{r}
reliance <- Equity$new(exchange = "nse", symbol = "RELIANCE")

answer <- reliance$place_order(
  transaction_type = "buy",
  order_type = "limit",
  quantity = 1,
  product = "cnc",
  price = 1000,
  dry_run = TRUE
)
print(answer)

answer <- reliance$place_order(
  transaction_type = "buy",
  order_type = "limit",
  quantity = 1,
  product = "cnc",
  price_reference = list(
    kind = "offer_level",
    level = 1
  ),
  dry_run = TRUE
)
print(answer)
```

When the Python version of these two calls was run against a local UBI on Saturday 2026-09-26, with the market closed and the day before UBI began holding plain limit orders, the answers showed three things worth knowing. The plain limit order's dry run may answer differently today.

- UBI chose a different broker for each request, `shoonya` and then `stoxkart`, because its selector takes turns among the brokers that can take the order.
- The lower-case `"buy"`, `"limit"` and `"cnc"` reached each broker in that broker's own spelling: `trantype: B`, `prctyp: LMT` and `prd: C` for Shoonya, and `DELIVERY` for Stoxkart.
- The price reference was resolved to `1226.0`, which was the one level on the offer side of the book in the quote captured the same afternoon.

Each dry-run answer held the `broker`, `dry_run` set to true, the `instrument_id`, an `intent_id`, the `request` with the broker's `method`, `url` and form or JSON body, an empty `skipped` list, a null `tag` and the preparation time under `timing_ms`. In R these arrive as a named list, so the resolved price of the second call is `answer[["request"]][["json"]][["price"]]`.

#### Returns

The method returns a named list, which is UBI's answer unchanged. The table below lists its elements; which of them appear depends on whether the order was sent, was a dry run, or is held inside UBI's order engine.

| Element | Type | Description |
|---|---|---|
| `broker` | character or `NULL` | The broker UBI chose. It is `NULL` for a held limit order or a synthetic order that is waiting for a price or a time. |
| `instrument_id` | character | The instrument the order was for. |
| `order_id` | character or `NULL` | The broker's id for the order, which `modify_order()` and `cancel_order()` take. It is `NULL` unless the outcome is `accepted`. |
| `outcome` | character | `accepted`, `rejected` or `unknown`; `partial` for a type that sends several orders when only some were accepted, which UBI answers with HTTP 207 and the package returns rather than signals; or `armed` for an order the engine is holding. Absent on a dry run. |
| `status_message` | character or `NULL` | Why the outcome is not `accepted`. |
| `broker_response` | named list, character or `NULL` | The broker's own answer. |
| `dry_run` | logical | `TRUE`, on a dry run only. |
| `request` | named list | On a dry run only: the method, the URL and the body the broker would have received. |
| `skipped` | list | Each broker UBI passed over before the one it chose, with its reason. |
| `tag` | character or `NULL` | The tag the order carried. |
| `timing_ms` | named list | How long UBI spent preparing the order, and how long the broker took. |
| `intent_id` | character | The id of the order's intent in UBI's order engine, on every answer. [`Account$intent()`](guide-account.html#intent) reads the answer again by it. |
| `parent_id` | character | Present for an order the engine recorded as a parent, including a held limit order. Keep it: it is the only handle on an order that has not reached a broker yet. |
| `legs` | list | Present for the types that send several orders at once, such as `freeze_slicer` and `ladder`, one entry per order with its plan `path`, `instrument_id`, `outcome`, `order_id` and `status_message`. |

#### Errors

The table below lists the conditions the method signals. Each one can be caught by its class name with `tryCatch()`, for example `tryCatch(..., ConflictError = function(error) ...)`.

| Condition | When |
|---|---|
| `BadRequestError` | A field is invalid, the price fields do not fit the order type, or a synthetic order's own settings are wrong. |
| `LossLockoutError` | The day's loss is past UBI's daily loss limit. |
| `NotFoundError` | No broker has a mapping for this instrument today. |
| `ConflictError` | A `quantity_reference` asked to reduce or close a position that is not held, a market order run as a marketable limit found nobody on the other side of the book or no fresh quote, a reduce-only order would not reduce the position, or the engine read the order too late or had already started it before a restart. |
| `OrderRejectedError` | The broker refused the order. Its answer is in the condition's `detail` field, not in its message. |
| `RateLimitError` | The broker's daily order cap has no room for this order. |
| `ServiceUnavailableError` | No broker could take the order, the order engine is not running, or a `price_reference` could not be resolved, for example because the book is empty. |
| `OrderOutcomeUnknownError` | The order was sent but its outcome is unknown. Read [`orders`](#orders), or [`Account$intent()`](guide-account.html#intent) with the `intent_id` in the condition's `detail`, before sending it again, or you may place it twice. |
| `UnifiedBrokerInterfaceError` | Any other failure reported by, or on the way to, UBI. |

Each of these is described, with what to do next, on the [Errors](guide-errors.html) page.

::: {.alert .alert-info}
**Under the hood.**

The method builds a named list from its arguments and sends it as JSON with the shared client's `post()` method. The six fields that are always present come first, and each optional field is added only when it is not `NULL`. For the price-reference example above, the body was the one below.

```json
{
  "instrument_id": "3f92570a-9924-5bf5-9f9d-e006cd9f4202",
  "transaction_type": "buy",
  "order_type": "limit",
  "product": "cnc",
  "after_market": false,
  "dry_run": true,
  "quantity": 1,
  "price_reference": {"kind": "offer_level", "level": 1}
}
```

Nothing else is sent: one call is one request. UBI's route is documented under [Place an order](https://pramodathani.github.io/unified_broker_interface/rest-api/orders/#place-an-order).
:::

## modify_order

The method's signature is shown below. It places orders, and it sends `PUT /api/orders/modify`.

```r
modify_order(
  order_id = NULL,
  quantity = NULL,
  price = NULL,
  trigger_price = NULL,
  order_type = NULL,
  validity = NULL,
  disclosed_quantity = NULL,
  broker = NULL,
  dry_run = FALSE,
  parent_id = NULL,
  part = NULL
)
```

This method changes one order that is still waiting in the market. Give at least one field to change; every field left as `NULL` keeps the value the order already has. UBI finds the order by its id in the brokers' order books, so the method does not check that the order belongs to the instrument you called it on.

UBI's order books are copies that its own collectors refresh every few seconds, so an order placed a moment ago is not in them yet, and changing it signals `NotFoundError`. In a live test on 2026-09-20 an order took 2.0 seconds to appear. Wait until the order shows up in [`orders`](#orders) before you change it.

An order that is a leg of one of UBI's synthetic orders is handed to the order engine, which lets the order type carry on from the change: a trailing stop trails from the trigger you set, and a chaser steps on from the price you set. Only such a leg's `price`, `trigger_price` and `quantity` can change, and anything else signals `ConflictError`.

An order the engine is still holding, such as a plain limit order waiting for the book to reach its price, has no broker order id yet. Name it by `parent_id` instead of `order_id`. Only its `price` and `quantity` can change, nothing is sent to a broker, and the order is later sent at the new terms.

#### Parameters

The table below lists the method's parameters.

| Name | Type | Required | Default | Description |
|---|---|:---:|---|---|
| `order_id` | character or `NULL` | one of the two | `NULL` | The id the broker gave the order, as `place_order()` returned it. |
| `quantity` | integer or `NULL` | no | `NULL` | The new total quantity in units, counting what has already filled. |
| `price` | numeric or `NULL` | no | `NULL` | The new limit price in rupees. |
| `trigger_price` | numeric or `NULL` | no | `NULL` | The new trigger price in rupees. |
| `order_type` | character or `NULL` | no | `NULL` | The new order type: `market`, `limit`, `sl` or `sl-m`. |
| `validity` | character or `NULL` | no | `NULL` | The new validity: `day` or `ioc`. |
| `disclosed_quantity` | integer or `NULL` | no | `NULL` | The new quantity to show on the exchange. |
| `broker` | character or `NULL` | no | `NULL` | The broker holding the order. It is needed only after a `ConflictError` that says two brokers share the id. |
| `dry_run` | logical | no | `FALSE` | `TRUE` has UBI build the broker's request and return it without sending it. |
| `parent_id` | character or `NULL` | one of the two | `NULL` | The `parent_id` of an order the engine is still holding, or of the plan that holds `part`, as `place_order()` returned it. |
| `part` | character or `NULL` | no | `NULL` | With `parent_id`, the path of a part of a `plan` order that has not been sent, such as `root.each_fill.children.0` for a bracket's stop. Its `price`, `trigger_price` and `quantity` can change, and it keeps them until its turn comes. |

#### Example

The first example below changes a held limit order by its `parent_id`. No output is shown for it, because it would change a real order.

```{r}
held <- reliance$buy_at_limit_price(price = 1000, quantity = 1, product = "cnc")
reliance$modify_order(parent_id = held[["parent_id"]], price = 995)
```

The second example sends a limit order to the broker at once, with the `simple` type, waits for it to appear and then lowers its price. No output is shown for it either, for the same reason.

```{r}
answer <- reliance$place_order(
  transaction_type = "buy",
  order_type = "limit",
  quantity = 1,
  product = "cnc",
  price = 1000,
  after_market = TRUE,
  synthetic = list(type = "simple")
)
order_id <- answer[["order_id"]]

for (attempt in 1:30) {
  frame <- reliance$open_orders
  if (!is.null(frame) && order_id %in% frame$order_id) {
    break
  }
  Sys.sleep(2)
}

reliance$modify_order(order_id = order_id, price = 995, broker = answer[["broker"]])
```

#### Returns

The method returns a named list with `broker`, `order_id`, `instrument_id`, `status_before_modify`, `outcome`, `status_message`, `broker_response` and `timing_ms`, plus `parent_id` and `synthetic_type` for a leg of a synthetic order. On a dry run it holds `dry_run` and the `request` UBI would have sent instead. A held order answers with `parent_id`, `synthetic_type`, `held` set to `TRUE`, the new `price` and `quantity`, and an `outcome` of `accepted`.

#### Errors

The table below lists the conditions the method signals.

| Condition | When |
|---|---|
| `BadRequestError` | No field was given to change, or a field is invalid or is one this broker cannot change. |
| `NotFoundError` | No broker's order book holds this order id, which is also what a very new order gives, or the engine holds no parent with this `parent_id`. |
| `ConflictError` | The order is already complete, cancelled, rejected or expired; two brokers hold the same id and the condition's `detail` lists them under `brokers`; a leg of a synthetic order was asked to change a field other than its price, trigger price or quantity; or a held order has already been sent, when the `detail` names its `broker` and `order_id`. |
| `OrderRejectedError` | The broker refused the change. |
| `ServiceUnavailableError` | The broker's order rate budget was full, so the change was not sent. It is worth retrying a moment later. |
| `OrderOutcomeUnknownError` | The change was sent but its outcome is unknown. |
| `UnifiedBrokerInterfaceError` | Any other failure reported by, or on the way to, UBI. |

::: {.alert .alert-info}
**Under the hood.**

The body holds `dry_run`, plus `order_id` or `parent_id` and each field that is not `NULL`, and it is sent with the shared client's `put()` method. UBI's route is documented under [Modify an order](https://pramodathani.github.io/unified_broker_interface/rest-api/orders/#modify-an-order).
:::

## cancel_order

The method's signature is shown below. It places orders, and it sends `DELETE /api/orders/cancel`.

```r
cancel_order(order_id, broker = NULL, dry_run = FALSE)
```

This method cancels one order that is still waiting in the market. Like `modify_order()`, it finds the order by id across every broker's order book, does not check which instrument it belongs to, and cannot see an order placed in the last few seconds.

An order that is a leg of one of UBI's synthetic orders is cancelled through the order engine, so the order type knows about it, but the synthetic order itself carries on. [`cancel_parent()`](#cancel_parent) stops a synthetic order, and it is also how an order the engine is still holding is cancelled, because such an order has no broker order id. Since 2026-10-05 a leg cancelled this way stays cancelled: a bracket's stop is not placed again, and the leg's unfilled quantity comes off what its part trades, so a later fill of the entry is protected for the smaller quantity only.

#### Parameters

The table below lists the method's parameters.

| Name | Type | Required | Default | Description |
|---|---|:---:|---|---|
| `order_id` | character | yes | | The id the broker gave the order. |
| `broker` | character or `NULL` | no | `NULL` | The broker holding the order, needed only when two brokers share the id. |
| `dry_run` | logical | no | `FALSE` | `TRUE` has UBI build the broker's request and return it without sending it. |

#### Example

The example below cancels an order, naming the broker from the answer that placed it.

```{r}
reliance$cancel_order(order_id, broker = answer[["broker"]])
```

#### Returns

The method returns a named list with `broker`, `order_id`, `status_before_cancel`, `outcome`, `status_message`, `broker_response` and `timing_ms`, plus `parent_id` and `synthetic_type` for a leg of a synthetic order, or, on a dry run, `dry_run` and the `request` UBI would have sent.

#### Errors

The table below lists the conditions the method signals.

| Condition | When |
|---|---|
| `BadRequestError` | The order id, the broker or the dry run flag is malformed. |
| `NotFoundError` | No broker's order book holds this order id. |
| `ConflictError` | The order is already complete, cancelled, rejected or expired, or two brokers hold the same id. |
| `OrderRejectedError` | The broker refused the cancellation. |
| `ServiceUnavailableError` | The broker's order rate budget was full, so the cancellation was not sent. |
| `OrderOutcomeUnknownError` | The cancellation was sent but its outcome is unknown. |
| `UnifiedBrokerInterfaceError` | Any other failure reported by, or on the way to, UBI. |

::: {.alert .alert-info}
**Under the hood.**

The body holds `order_id`, `dry_run` and, when given, `broker`, and it is sent with the shared client's `delete()` method. UBI's route is documented under [Cancel an order](https://pramodathani.github.io/unified_broker_interface/rest-api/orders/#cancel-an-order).
:::

## cancel_open_orders

The method's signature is shown below. It places orders: it sends `GET /api/orders/parents` and `GET /api/orders/details`, then one `DELETE /api/orders/cancel` per parent and one more `DELETE /api/orders/cancel` with a list of orders for the rest.

```r
cancel_open_orders()
```

This method cancels every order in this instrument that is still waiting, whether it rests at a broker or is held in UBI's order engine. The numbered steps below are what it does.

1. It reads this instrument's open [`parents`](#parents) and cancels each with [`cancel_parent()`](#cancel_parent). The parents go first, because a synthetic order left running could place a new order after its old ones had been cancelled.
2. It reads [`open_orders`](#open_orders) and leaves out every order whose `engine_parent_id` is a parent the engine took a cancel for, because the engine has already dealt with it. An order whose parent's cancel failed is kept, so it is still cancelled.
3. It cancels every remaining order in one request, naming the broker from each order's row, using the list form of UBI's cancel route.

Every parent and order is attempted even when an earlier one fails, and a failure is reported in the returned data frame rather than signalled, so one order that can no longer be cancelled does not leave the rest open. A parent whose cancel a broker refused for one leg is reported as not cancelled, because that leg may still be live.

It has no `dry_run` argument. To see what it would cancel, read `parents` and `open_orders` yourself first.

#### Parameters

The method takes no parameters.

#### Example

The example below cancels everything waiting in RELIANCE and prints what happened. No real output is shown, because running the method cancels real orders. As an illustration, a run that found a held limit order, one broker order still open and one broker order that had filled a moment before would give three rows: the held order's parent with `cancelled` set to `TRUE`, the open broker order with `cancelled` set to `TRUE`, and the filled one with `cancelled` set to `FALSE` and an `error` of `HTTP 409: the order is already COMPLETE`, which is UBI's message for an order that is already finished.

```{r}
outcome <- reliance$cancel_open_orders()
print(outcome)
```

#### Returns

The method returns a `data.frame` with one row per parent and per order, or `NULL` when nothing in this instrument is waiting. The table below lists its columns.

| Column | Type | Description |
|---|---|---|
| `parent_id` | character or `NA` | The parent cancelled, or `NA` for an order cancelled on its own. |
| `order_id` | character or `NA` | The broker's id for the order, or `NA` for a parent. |
| `broker` | character or `NA` | The broker holding the order, or `NA` for a parent. |
| `cancelled` | logical | Whether the cancellation was accepted. |
| `error` | character or `NA` | Why it was not: the condition's class name and message for a parent, or the HTTP status and UBI's message for an order. `NA` when it was. |

#### Errors

The table below lists the conditions the method signals.

| Condition | When |
|---|---|
| `BrokerError` | No broker's order book could be read. |
| `ServiceUnavailableError` | UBI's order book document is missing or too old to serve, or UBI's parents could not be read. |
| `UnifiedBrokerInterfaceError` | The order book or the parents could not be read for any other reason, or UBI refused the list of cancels as a whole. A failure to cancel one order or parent is reported in the data frame instead. |

## Reading the order book

UBI serves the whole account's order book and trade book, across every broker, and has no route for one instrument. So each active binding below reads the whole book, one request per access, and keeps the rows whose `instrument_id` matches this instrument. The book is not merged across brokers: one order at one broker is one row, and the same instrument traded at two brokers gives a row from each. A row whose `instrument_id` UBI could not work out is invisible here, because there is no other field that names the instrument reliably.

The table below shows which statuses each active binding keeps. An order still waiting in the market is `PENDING` at some brokers and `OPEN` at others, which is why `open_orders` takes both.

| Active binding | Statuses kept |
|---|---|
| `orders` | all six |
| `open_orders` | `PENDING`, `OPEN` |
| `completed_orders` | `COMPLETE` |
| `rejected_orders` | `REJECTED` |
| `cancelled_orders` | `CANCELLED` |

A status with no active binding of its own, such as `EXPIRED`, is found by filtering the `status` column of `orders`. An order UBI's order engine is still holding has not reached a broker, so it is in none of these data frames; it is in [`parents`](#parents). Every active binding returns `NULL`, not an empty data frame, when no row matches, and each access sends a new request, so assign the data frame to a variable when you need it twice. The active bindings are read-only, and assigning to one signals an error.

### orders

This active binding reads `GET /api/orders/details`. It gives every one of today's orders in this instrument, whatever its status.

#### Example

The example below reads the orders and keeps the expired ones. When the Python version was run against a local UBI on 2026-09-26, the account had placed no RELIANCE order that day, so the property returned nothing, which in R is `NULL`.

```{r}
frame <- reliance$orders
print(frame)

if (!is.null(frame)) {
  expired <- frame[frame$status == "EXPIRED", ]
}
```

#### Returns

The active binding gives a `data.frame` with UBI's order fields, or `NULL` when this instrument has no orders today. The table lists the columns the package's documentation names; UBI's [order book](https://pramodathani.github.io/unified_broker_interface/rest-api/orders/#order-book) page lists every field.

| Column | Type | Description |
|---|---|---|
| `broker` | character | The broker holding the order. |
| `order_id` | character | The broker's id for the order. |
| `status` | character | `PENDING`, `OPEN`, `COMPLETE`, `CANCELLED`, `REJECTED` or `EXPIRED`. |
| `status_message` | character or `NA` | Why an order was rejected, in the words of whoever refused it. |
| `transaction_type` | character | `BUY` or `SELL`. |
| `product` | character | `CNC`, `MIS` or `NRML`. |
| `order_type` | character | `MARKET`, `LIMIT`, `SL` or `SL-M`. |
| `quantity` | numeric | The order's quantity in units. |
| `filled_quantity` | numeric | How much has filled. |
| `price` | numeric | The limit price. |
| `trigger_price` | numeric | The trigger price. |
| `average_price` | numeric | The average fill price. |
| `order_timestamp` | character | When the order was placed. |
| `engine_parent_id` | character or `NA` | The order engine parent that placed the order, or `NA` for an order placed elsewhere. |
| `leg_role` | character or `NA` | The order's role in that parent, such as `entry`, `stop`, `target` or `slice`. |
| `synthetic_type` | character or `NA` | The parent's order type, such as `bracket` or `virtual_limit`. |
| `intent_id` | character or `NA` | The intent the parent was placed for, as the place answer named it. |

A column that is null in every row comes back as a logical column of `NA` values rather than a character one.

#### Errors

The table below lists the conditions the active binding signals.

| Condition | When |
|---|---|
| `BrokerError` | No broker's order book could be read. |
| `ServiceUnavailableError` | UBI's order book document is missing or too old to serve. This means UBI's own background writer stopped, not that a broker is down. |
| `UnifiedBrokerInterfaceError` | Any other failure reported by, or on the way to, UBI. |

### open_orders

This active binding reads `GET /api/orders/details`. It gives today's orders in this instrument that are still waiting in the market, which are the only ones `modify_order()` and `cancel_order()` will accept. It returns a data frame shaped like [`orders`](#orders), or `NULL`, and signals the same conditions.

### completed_orders

This active binding reads `GET /api/orders/details`. It gives today's orders in this instrument that filled in full. It returns a data frame shaped like [`orders`](#orders), or `NULL`, and signals the same conditions.

### rejected_orders

This active binding reads `GET /api/orders/details`. It gives today's orders in this instrument that a broker or the exchange refused. The `status_message` column holds the reason, which is the first place to look after an `accepted` order that never filled. It returns a data frame shaped like [`orders`](#orders), or `NULL`, and signals the same conditions.

### cancelled_orders

This active binding reads `GET /api/orders/details`. It gives today's orders in this instrument that were cancelled. It returns a data frame shaped like [`orders`](#orders), or `NULL`, and signals the same conditions.

### trades

This active binding reads `GET /api/orders/trades`. It gives today's trades in this instrument. One order can fill in several trades, and each trade names the order it came from, so `trades` is where to look for the prices an order actually filled at.

#### Returns

The active binding gives a `data.frame` with UBI's trade fields, or `NULL` when this instrument has no trades today. The columns the package's documentation names are listed below; UBI's [trade book](https://pramodathani.github.io/unified_broker_interface/rest-api/orders/#trade-book) page lists every field.

| Column | Type | Description |
|---|---|---|
| `broker` | character | The broker the trade happened at. |
| `trade_id` | character | The id of the trade. |
| `order_id` | character | The order the trade filled. |
| `transaction_type` | character | `BUY` or `SELL`. |
| `product` | character | `CNC`, `MIS` or `NRML`. |
| `quantity` | numeric | The quantity traded. |
| `price` | numeric | The price it traded at. |
| `value` | numeric | The quantity times the price. |
| `trade_timestamp` | character | When it traded. |
| `engine_parent_id`, `leg_role`, `synthetic_type`, `intent_id` | character or `NA` | The order engine parent the trade's order belongs to, as for an order, or `NA` for an order placed elsewhere. |

#### Errors

The table below lists the conditions the active binding signals.

| Condition | When |
|---|---|
| `BrokerError` | No broker's trade book could be read. |
| `ServiceUnavailableError` | UBI's trade book document is missing or too old to serve. |
| `UnifiedBrokerInterfaceError` | Any other failure reported by, or on the way to, UBI. |

## Reading the order engine's parents

A parent is one order the order engine was asked for, such as a bracket, a trailing stop or a held limit order, and its legs are the broker orders it placed. Every answer from the engine for such an order carries a `parent_id`, which the members below take. The classes on [Synthetic orders](guide-synthetic-orders.html) keep their own `parent_id` and call these members for you.

### cancel_parent

The method's signature is shown below. It places orders, and it sends `DELETE /api/orders/cancel`.

```r
cancel_parent(parent_id, part = NULL, dry_run = FALSE)
```

This method cancels one parent, with every leg it still has resting at a broker, so the parent places, moves and cancels nothing more. It is how a synthetic order is stopped and how a held limit order is cancelled. A position the parent has already opened is not closed.

When a broker refuses the cancel of one leg, or its outcome is unknown, UBI answers HTTP 207, which the package returns rather than signals, and the parent's `state` is `cancelling` rather than `cancelled`. The parent no longer acts, and becomes `cancelled` on its own once the broker reports that leg finished. Read `cancelled_legs` to see which leg may still be live, and call the method again to retry it.

#### Parameters

The table below lists the method's parameters.

| Name | Type | Required | Default | Description |
|---|---|:---:|---|---|
| `parent_id` | character | yes | | The `parent_id` that `place_order()` answered with. |
| `part` | character or `NULL` | no | `NULL` | The path of one part of a `plan` order to cancel, such as `root.each_fill.children.0`, as the parent's `parameters$parts` lists it. The rest of the plan carries on. `NULL` cancels the whole parent. |
| `dry_run` | logical | no | `FALSE` | `TRUE` has UBI say what would be cancelled, without cancelling anything. |

#### Example

The example below places a held limit order and cancels it again. No output is shown for it, because it places and cancels a real order.

```{r}
held <- reliance$buy_at_limit_price(price = 1000, quantity = 1, product = "cnc")
answer <- reliance$cancel_parent(held[["parent_id"]])
print(answer[["state"]])
```

#### Returns

The method returns a named list with `parent_id`, `synthetic_type`, `state`, `intent_id` and `cancelled_legs`, which holds one entry per leg with its `leg_id`, `broker`, `order_id`, `outcome` and `status_message`. A part answers instead with `parent_id`, `synthetic_type`, `part`, its `state`, `outcome`, `status_message`, `intent_id` and `orders`, where each order's `cancel_accepted` says whether its broker accepted the cancel.

#### Errors

The table below lists the conditions the method signals.

| Condition | When |
|---|---|
| `BadRequestError` | The parent id is malformed. |
| `NotFoundError` | The order engine holds no parent with this id, or the plan has no part at this path. |
| `ConflictError` | The parent or part has already finished, the part is kept whole and has not started, or the parent is not a plan and was given a part. |
| `ServiceUnavailableError` | The order engine is not running. |
| `OrderOutcomeUnknownError` | The engine did not answer in time. |
| `UnifiedBrokerInterfaceError` | Any other failure reported by, or on the way to, UBI. |

### parents

This active binding reads `GET /api/orders/parents`. It gives this instrument's parents that the order engine has not finished. It is the only way to see a parent that has placed nothing yet, such as a held limit order or an armed trigger. UBI lists every open parent in the account, so the active binding reads them all and keeps this instrument's; [`Account$parents`](guide-account.html#parents) gives the whole list.

#### Returns

The active binding gives a `data.frame` with one row per parent, or `NULL` when no parent in this instrument is open. The table lists its main columns. The columns that hold nested objects are list columns, so `frame$body[[1]]` is the first parent's body as a named list.

| Column | Type | Description |
|---|---|---|
| `parent_order_id` | character | The parent's id, which is the `parent_id` the other members take. |
| `synthetic_type` | character | `plan` for every type but `simple`, since UBI runs every other type as a plan of its preset; the type asked for is under `parameters` as `routed_from`. |
| `state` | character | `received`, `working` or `cancelling`; a parent placed before 2026-10-03 can also show `protecting`. The finished states, `completed`, `cancelled`, `rejected` and `failed`, do not appear here. |
| `instrument_id` | character | This instrument. |
| `body` | list column of named lists | The order body the parent was placed with. |
| `parameters` | list column of named lists | The type's settings, including the engine's own working values. |
| `legs` | list column of lists | One entry per broker order the parent placed. |

#### Errors

The table below lists the conditions the active binding signals.

| Condition | When |
|---|---|
| `ServiceUnavailableError` | UBI's parents could not be read. |
| `UnifiedBrokerInterfaceError` | Any other failure reported by, or on the way to, UBI. |

### parent

This method's signature is `parent(parent_id)`, and it reads `GET /api/orders/parents?parent_id=...`. It reads one parent, whether or not it has finished. UBI finds it by id alone, so the method does not check that it belongs to this instrument. It returns a named list with the same fields as a row of [`parents`](#parents), and signals `NotFoundError` when the engine holds no parent with this id.

### parent_orders

This method's signature is `parent_orders(parent_id)`, and it reads `GET /api/orders/details?parent_id=...`. It gives today's broker orders that one parent placed, using the order book's `parent_id` filter. It returns a data frame shaped like [`orders`](#orders), whose `leg_role` column says what each order was to the parent, or `NULL` when the parent has placed nothing the order book shows yet. It signals the same conditions as `orders`.

### parent_trades

This method's signature is `parent_trades(parent_id)`, and it reads `GET /api/orders/trades?parent_id=...`. It gives today's trades in the broker orders that one parent placed. It returns a data frame shaped like [`trades`](#trades), or `NULL` when none of the parent's orders has traded, and signals the same conditions as `trades`.
