---
title: "Holding for the long term"
---

```{r, include = FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

A holding is something you own outright and keep in your demat account, such as shares bought for delivery, as against a position, which a futures, options or intraday trade leaves open. The six members on this page read a holding and add to it or sell out of it. Positions have their own page, [Positions](guide-positions.html).

::: {.alert .alert-danger}
**These are real orders.**

`add_to_holdings()`, `reduce_holdings()` and `liquidate_holdings()` send market or limit orders through UBI to real brokers, with real money, and none of them takes a `dry_run` argument. To see what one would send, call [`place_order()`](guide-orders.html#place_order) with `product = "cnc"`, the same side, quantity and price, and `dry_run = TRUE`.
:::

The table below lists the six members.

| Kind | Member | Description |
|---|---|---|
| active binding | [`holdings`](#holdings) | The long-term holding of this instrument, merged across every broker. |
| active binding | [`holdings_value`](#holdings_value) | What the units held are worth at the moment. |
| active binding | [`holdings_pnl`](#holdings_pnl) | What the units held have made or lost. |
| places orders | [`add_to_holdings`](#add_to_holdings) | Buys more of this instrument to keep. |
| places orders | [`reduce_holdings`](#reduce_holdings) | Sells some of the units held, without selling more than are free. |
| places orders | [`liquidate_holdings`](#liquidate_holdings) | Sells every unit held that is free to sell. |

## Which classes can be held

Only instruments in UBI's cash segments can be reported as holdings, so only five classes carry these members. Each class has its own copy of them rather than sharing a base class, but the six behave the same way in all five. The table below shows which classes have them and why the rest do not.

| Class | Holdings members | Why |
|---|:---:|---|
| [`Equity`](../reference/Equity.html) | Yes | A share is the classic holding. |
| [`FixedIncome`](../reference/FixedIncome.html) | Yes | A bond or a sovereign gold bond is held in the demat account. |
| [`ExchangeTradedFund`](../reference/ExchangeTradedFund.html) | Yes | A fund's units trade and are held like shares. |
| [`InvestmentTrust`](../reference/InvestmentTrust.html) | Yes | A trust's units trade and are held like shares. |
| [`MutualFund`](../reference/MutualFund.html) | Yes | A scheme is subscribed to rather than traded, but holding it is what it is for. |
| Every futures and options class | No | A derivative leaves a position, not a holding. |
| Every index class | No | An index cannot be owned. |
| `Commodity`, `Currency` and their derivatives | No | UBI's cash segments exclude commodities and currencies, so they are never reported as holdings. |

The asset class pages, [Equities](asset-classes-equities.html), [Fixed income](asset-classes-fixed-income.html), [Funds](asset-classes-funds.html) and [Mutual funds](asset-classes-mutual-funds.html), describe each of these classes in full.

Two classes need care. No broker that serves quotes carries a cash bond or a mutual fund, so `FixedIncome` and `MutualFund` have no live price. Give the three order methods a `price`, which sends a limit order, rather than letting them send a market order into a book nobody quotes. When no price is given, these two classes send a real market order, because the `marketable_limit` UBI makes of a plain market order needs a quote and would always be refused with HTTP 409.

## Always cnc

Every order these members send uses the `cnc` product, and there is no argument to change it. `cnc` is the only product that buys into or sells out of a demat account. Selling a holding as `mis` would not sell your shares at all: it would open an intraday short position beside them, which the broker squares off before the close, so the mistake would cost money twice and leave the holding untouched. Leaving the argument out makes that impossible to do by accident.

The order methods send plain market or limit orders through [`buy_at_market_price()`](guide-price-wrappers.html#buy_at_market_price), [`buy_at_limit_price()`](guide-price-wrappers.html#buy_at_limit_price) and their sell twins. A limit order is held by UBI's order engine until the book reaches its price, except for `MutualFund` and `FixedIncome`, which pass `hold = FALSE` because nothing quotes them and a held order would never be sent. A market order is sent by UBI as a `marketable_limit`, a limit that follows the other side of the book for 30 seconds, except again for those two classes, which pass `as_marketable_limit = FALSE` for the same reason. [Order engine](architecture-order-engine.html#when-to-send-a-limit-order-at-once) explains why.

## Free units and pledged units

A holding's `collateral_quantity` is the part pledged to a broker as margin, and a broker will not let you sell it until the pledge is released. So `reduce_holdings()` and `liquidate_holdings()` work on the free units, which are the whole `quantity` minus `collateral_quantity`, and refuse before sending anything when the free units cannot cover the sale. The flowchart below shows the check both methods make.

<pre class="mermaid">
flowchart TD
    A["reduce_holdings() or liquidate_holdings()"] --&gt; B["read the holding once"]
    B --&gt; C{"held at all?"}
    C -- "no" --&gt; E1["HoldingError, nothing is sent"]
    C -- "yes" --&gt; D["free units = quantity minus collateral_quantity"]
    D --&gt; F{"which method?"}
    F -- "reduce_holdings()" --&gt; G{"quantity asked for&lt;br/&gt;more than the free units?"}
    G -- "yes" --&gt; E1
    G -- "no" --&gt; H["sell the quantity asked for, as cnc"]
    F -- "liquidate_holdings()" --&gt; I{"any free units?"}
    I -- "no, all pledged" --&gt; E1
    I -- "yes" --&gt; J["sell every free unit, as cnc"]
</pre>

`holdings_value` and `holdings_pnl` still count pledged units, because a pledged share is still owned.

## holdings

`holdings` is an active binding, and each read sends `GET /api/portfolio/holdings`.

This active binding gives the long-term holding of this instrument, merged across every broker. UBI serves the whole account's holdings and has no route for one instrument, so each access sends one request and picks this instrument's row. It matches on `instrument_id` first, and then on `symbol`, because UBI merges a holding held on two exchanges into one row under whichever broker's row arrived first, and that row may carry the other exchange's id.

Unlike the order and position readers, it returns a single named list rather than a data frame, because an instrument has at most one holdings row.

#### Example

The example below reads the holding of two shares.

```{r}
reliance <- Equity$new(exchange = "nse", symbol = "RELIANCE")
print(reliance$holdings)

infosys <- Equity$new(exchange = "nse", symbol = "INFY")
str(infosys$holdings)
```

When the Python version of the first half ran against a local UBI on 2026-09-26, no RELIANCE shares were held, so the holding was empty; in R, `reliance$holdings` is then `NULL`. For the second half, UBI's [holdings](https://pramodathani.github.io/unified_broker_interface/rest-api/portfolio/#holdings) page shows an illustrative INFY row, which the table below lists. With that row, `infosys$holdings` is a named list holding these values, and its `pnl` element is itself a named list.

| Name | Value in the illustrative row |
|---|---|
| `instrument_id` | `11111111-1111-5111-8111-000000000002` |
| `isin` | `INE009A01021` |
| `symbol` | `INFY` |
| `exchange` | `nse` |
| `segment` | `nse_equities` |
| `quantity` | 20 |
| `average_price` | 1400 |
| `invested_value` | 28000 |
| `last_price` | 1521.4 |
| `current_value` | 30428 |
| `pnl` | `unrealized` 2428, `day_change` 13.25 and `day_change_percentage` 0.88 |
| `collateral_quantity` | 0 |

#### Returns

A named list with the elements below, or `NULL` when no broker holds this instrument.

| Name | Type | Description |
|---|---|---|
| `instrument_id` | character | The instrument. |
| `isin` | character or `NULL` | The ISIN, when a broker sent one. |
| `symbol` | character | The symbol. |
| `exchange` | character | The exchange. |
| `segment` | character | The exchange-prefixed segment. |
| `quantity` | numeric | Everything held, across every broker, including pledged units. |
| `average_price` | numeric | What was paid per unit on average. |
| `invested_value` | numeric | What was paid in all. |
| `last_price` | numeric or `NULL` | The price used to value the holding. |
| `current_value` | numeric or `NULL` | `quantity` times `last_price`. |
| `pnl` | named list | `unrealized`, `day_change` and `day_change_percentage`. |
| `collateral_quantity` | numeric | The units pledged as collateral, which cannot be sold. |

#### Errors

| Condition | When |
|---|---|
| `ServiceUnavailableError` | UBI's holdings document is missing or too old to serve. |
| `BrokerError` | No broker's holdings could be read. |
| `UnifiedBrokerInterfaceError` | Any other failure reported by, or on the way to, UBI. |

## holdings_value

`holdings_value` is an active binding, and each read sends `GET /api/portfolio/holdings`.

This active binding says what the units held are worth, in rupees. UBI prices a holding itself, so this reads the row's `current_value` rather than working it out, which is the opposite of [`positions_value`](guide-positions.html#positions_value). It counts every unit held, including pledged ones.

#### Example

The example below reads the value of the INFY holding.

```{r}
print(infosys$holdings_value)
```

With the same illustrative INFY row, it prints 30428.

#### Returns

A numeric value in rupees, or `NULL` when this instrument is not held. It can also be `NULL` when the holding is held only at a broker that sends no price and there is no usable quote.

#### Errors

It signals the same conditions as [`holdings`](#holdings).

## holdings_pnl

`holdings_pnl` is an active binding, and each read sends `GET /api/portfolio/holdings`.

This active binding says what the units held have made or lost. Its named list is not shaped like a position's, and the table below shows the difference, because the two are easy to confuse.

| Name | `holdings_pnl` | [`positions_pnl`](guide-positions.html#positions_pnl) |
|---|:---:|:---:|
| `unrealized` | Yes | Yes |
| `day_change` | Yes | No |
| `day_change_percentage` | Yes | No |
| `realized` | No | Yes |
| `total` | No | Yes |

Only `unrealized` means the same thing in both: what is still riding on what you hold, against what you paid. A holding has no realised figure, because selling a share removes it from the holding rather than booking a profit against it.

#### Example

The example below reads the profit and loss of the INFY holding.

```{r}
str(infosys$holdings_pnl)
```

With the same illustrative INFY row, it is a named list with `unrealized` 2428, `day_change` 13.25 and `day_change_percentage` 0.88.

#### Returns

A named list with `unrealized` in rupees against what was paid, and `day_change` in rupees and `day_change_percentage` in per cent since the previous close, or `NULL` when this instrument is not held.

#### Errors

It signals the same conditions as [`holdings`](#holdings).

## add_to_holdings

`add_to_holdings(quantity, price = NULL, validity = NULL, after_market = FALSE, tag = NULL)` places orders, and each call sends `POST /api/orders/place`.

This method buys more of this instrument to keep, always as `cnc`. It reads nothing first, because you can buy whether or not you already hold any, and neither the package nor UBI checks your funds before sending; the broker does.

#### Parameters

| Name | Type | Required | Default | Description |
|---|---|:---:|---|---|
| `quantity` | integer | yes | | The number of units to buy. |
| `price` | numeric or `NULL` | no | `NULL` | A limit price in rupees, or `NULL` for a market order. |
| `validity` | character or `NULL` | no | `NULL` | `"day"` or `"ioc"`. UBI uses `"day"` when it is `NULL`. |
| `after_market` | logical | no | `FALSE` | `TRUE` sends an after-market order. |
| `tag` | character or `NULL` | no | `NULL` | A label of up to twenty letters and digits. |

#### Example

The example below buys ten INFY shares at no more than 1500 rupees each, and queues the order for the next session.

```{r}
infosys$add_to_holdings(quantity = 10, price = 1500.0, after_market = TRUE)
```

#### Returns

The named list that [`place_order()`](guide-orders.html#place_order) returns.

#### Errors

| Condition | When |
|---|---|
| `UnifiedBrokerInterfaceError` | Any failure reported by, or on the way to, UBI, including every condition [`place_order()`](guide-orders.html#place_order) can signal. |

## reduce_holdings

`reduce_holdings(quantity, price = NULL, validity = NULL, after_market = FALSE, tag = NULL)` places orders, and each call sends `GET /api/portfolio/holdings` and then `POST /api/orders/place`.

This method sells some of the units held, as `cnc`. It reads the holding once and refuses, before sending anything, when the quantity asked for is more than the free units.

#### Parameters

| Name | Type | Required | Default | Description |
|---|---|:---:|---|---|
| `quantity` | integer | yes | | The number of units to sell. |
| `price` | numeric or `NULL` | no | `NULL` | A limit price in rupees, or `NULL` for a market order. |
| `validity` | character or `NULL` | no | `NULL` | `"day"` or `"ioc"`. |
| `after_market` | logical | no | `FALSE` | `TRUE` sends an after-market order. |
| `tag` | character or `NULL` | no | `NULL` | A label of up to twenty letters and digits. |

#### Example

The example below sells five of the INFY shares at no less than 1550 rupees each.

```{r}
infosys$reduce_holdings(quantity = 5, price = 1550.0)
```

#### Returns

The named list that [`place_order()`](guide-orders.html#place_order) returns.

#### Errors

| Condition | When |
|---|---|
| `HoldingError` | This instrument is not held, or the quantity is more than the free units. |
| `UnifiedBrokerInterfaceError` | Any failure reported by, or on the way to, UBI. |

## liquidate_holdings

`liquidate_holdings(price = NULL, validity = NULL, after_market = FALSE, tag = NULL)` places orders, and each call sends `GET /api/portfolio/holdings` and then `POST /api/orders/place`.

This method sells every unit held that is free to sell, as `cnc`. Pledged units are left alone, so it empties the holding only when nothing is pledged.

#### Parameters

| Name | Type | Required | Default | Description |
|---|---|:---:|---|---|
| `price` | numeric or `NULL` | no | `NULL` | A limit price in rupees, or `NULL` for a market order. |
| `validity` | character or `NULL` | no | `NULL` | `"day"` or `"ioc"`. |
| `after_market` | logical | no | `FALSE` | `TRUE` sends an after-market order. |
| `tag` | character or `NULL` | no | `NULL` | A label of up to twenty letters and digits. |

#### Example

The example below sells every free INFY share at market.

```{r}
infosys$liquidate_holdings()
```

#### Returns

The named list that [`place_order()`](guide-orders.html#place_order) returns.

#### Errors

| Condition | When |
|---|---|
| `HoldingError` | This instrument is not held, or every unit held is pledged as collateral. |
| `UnifiedBrokerInterfaceError` | Any failure reported by, or on the way to, UBI. |

## HoldingError

`HoldingError` is the package's own condition for a holding that cannot be changed as asked. It is signalled before anything is sent, and its parent class is `InstrumentError`, as the table `ASSETS_ERROR_PARENTS` in `R/assets_exceptions.R` records. The table below lists its three messages as `Equity` words them, with `<...>` marking a value filled in. The other four classes say `units` where `Equity` says `shares`, and in all five classes the first message ends with the instrument's `format()` text.

| Message | Signalled by |
|---|---|
| `No <symbol> shares are held, so there is nothing to sell` | `reduce_holdings()`, `liquidate_holdings()` |
| `<free> of the <quantity> <symbol> shares held are free to sell, because <pledged> are pledged as collateral, so <asked> cannot be sold` | `reduce_holdings()` |
| `All <quantity> <symbol> shares held are pledged as collateral, so none can be sold` | `liquidate_holdings()` |

The example below catches it, so a sale that is too large is reported rather than stopping the script.

```{r}
tryCatch(
  infosys$reduce_holdings(quantity = 1000000, price = 1550.0),
  HoldingError = function(error) {
    cat("Nothing was sold:", conditionMessage(error), "\n")
  }
)
```

The [Errors](guide-errors.html#holdingerror) page lists `HoldingError` with the rest of the package's error conditions.
