---
title: "Asset baskets"
---

```{r, include = FALSE}
knitr::opts_chunk$set(eval = FALSE)
```

Everything else in the package describes one instrument. An asset basket describes a group of them: a portfolio, a watchlist, an index, or what an exchange traded fund or a mutual fund holds. Every basket reads what it needs about all its members in one list request to UBI, and every basket can be analysed exactly like an instrument, from moving averages to the Sharpe ratio.

::: {.alert .alert-danger}
**Two members place real orders.**

`Portfolio$place_orders()` and `Portfolio$rebalance()` send one real order per instrument, in one request, with real money. UBI sends each market order as a limit that follows the other side of the book and is cancelled after 30 seconds, so a member with an empty book or no fresh quote gets HTTP 409 in its row and nothing is sent for it; pass `as_marketable_limit = FALSE` to send real market orders. The orders are placed in parallel and not as one unit, so some can be accepted while others are refused. Run them with `dry_run = TRUE` first and read the data frame they return.
:::

## The classes

The baskets are one base class and five kinds of basket, each in its own file, `R/asset_baskets_*.R`. The diagram below shows how they relate and which instrument classes link to them.

<pre class="mermaid">
classDiagram
    direction LR
    class AssetBasket {
        members
        weights
        prices()
        sharpe_ratio()
    }
    AssetBasket &lt;|-- Portfolio
    AssetBasket &lt;|-- Watchlist
    AssetBasket &lt;|-- Index
    AssetBasket &lt;|-- ExchangeTradedFundConstituents
    AssetBasket &lt;|-- MutualFundConstituents
    NonTradeableInstrument ..&gt; Index : constituents
    ExchangeTradedFund ..&gt; ExchangeTradedFundConstituents : constituents
    MutualFund ..&gt; MutualFundConstituents : constituents
</pre>

The table below says what each kind of basket is for.

| Class | Members carry | What it is for |
|---|---|---|
| [`AssetBasket`](../reference/AssetBasket.html) | a weight, or nothing | The shared base: live prices, weights, correlation, risk, and every analysis method |
| [`Portfolio`](../reference/Portfolio.html) | a quantity and an average price | What you hold: value, profit and loss, buying the whole basket, rebalancing |
| [`Watchlist`](../reference/Watchlist.html) | nothing | Instruments followed together, each counting equally |
| [`Index`](../reference/Index.html) | a weight | A weighted index, published or your own, and turning it into a portfolio |
| [`ExchangeTradedFundConstituents`](../reference/ExchangeTradedFundConstituents.html) | a weight | What an ETF holds, compared with the ETF's own price |
| [`MutualFundConstituents`](../reference/MutualFundConstituents.html) | a weight | What a mutual fund holds, the only way to measure one since UBI has no NAV |

## An index or a fund is two things

NIFTY is an instrument with its own price, which UBI quotes. It is also a basket of fifty shares with weights. The package keeps these apart rather than merging them into one class, because a merged class could not say whether `prices()` meant the official price or the price rebuilt from the members. The instrument keeps its price, and its `constituents` active binding returns the basket. The table below compares the two for each kind of instrument that has a basket.

| Class | `instrument$sharpe_ratio(days = 365)` | `instrument$constituents$sharpe_ratio(days = 365)` |
|---|---|---|
| `EquityIndex` (NIFTY) | Works on the official index candles | Works on the rebuilt basket; comparing the two shows how well the stored weights reproduce the index |
| `ExchangeTradedFund` (NIFTYBEES) | Works on the fund's traded price | Works on its holdings; `fund$tracking_error(benchmark = fund$constituents)` shows how closely the fund follows them |
| `MutualFund` | Returns `NULL`, because UBI stores no candles or NAV for mutual funds | The only way to measure a mutual fund |

`holdings` on a fund still means the units of the fund your account holds, which is why the fund's own contents are called `constituents`.

## Where the members come from

UBI stores nothing about what an index or a fund holds, so baskets live in this project's MongoDB, in the `asset_baskets` collection, which the package reaches through the mongolite package. You supply the members, by hand in R or from a CSV file. Each stored document is one version of one basket, dated by the day it takes effect, because an index is rebalanced and a fund's holdings change every month. The flowchart below shows the two ways in and the way back out.

<pre class="mermaid">
flowchart LR
    CSV["CSV file&lt;br/&gt;symbol, weight"] --&gt; IMP["BasketCsvImporter"]
    R["R&lt;br/&gt;list of BasketMember"] --&gt; B["a basket object"]
    IMP --&gt;|"one POST /api/instruments/details"| B
    B --&gt;|"BasketStore$save()"| DB[("MongoDB&lt;br/&gt;asset_baskets")]
    DB --&gt;|"BasketStore$load()&lt;br/&gt;or instrument$constituents"| B2["the basket again"]
</pre>

## AssetBasket

The constructor of [`AssetBasket`](../reference/AssetBasket.html) has the signature below.

```r
AssetBasket$new(name, members, linked_instrument = NULL, unmapped_weight = 0.0, unified_broker_interface = NULL)
```

The base class holds a list of [`BasketMember`](../reference/BasketMember.html) objects, each an instrument with a weight, a quantity, or neither; a member is made with `BasketMember$new(instrument, weight, quantity, average_price)`. Every member must be a different instrument, and either every member has a weight or none does; without weights the members count equally. `unmapped_weight` is the share of a fund held in things UBI cannot price, such as cash.

The table below lists the members every basket has. Where the Python library returns a pandas Series indexed by member label, the R package returns a named numeric vector whose names are the labels, and where Python returns a DataFrame indexed by label, R returns a data frame whose row names are the labels.

| Kind | Member | Description | UBI route |
|---|---|---|---|
| active binding | `instruments`, `labels`, `size` | What the basket holds; a label reads like `nse:INFY` | none |
| active binding | `weights` | The weights normalised to sum to 1, or equal weights, as a named numeric vector | none |
| active binding | `last_prices`, `ohlc`, `quotes` | One row per member, with an `error` column for a member UBI has no price for | `POST` `/ltp`, `/ohlc`, `/quote` |
| active binding | `day_change_percent` | The weighted day move of the whole basket | `POST` `/ohlc` |
| active binding | `advancers`, `decliners`, `breadth` | How many members are up, down, unchanged or unpriced today; `breadth` is a named list | `POST` `/ohlc` |
| active binding | `exposure_by_segment`, `exposure_by_exchange` | Where the weight sits | none |
| active binding | `concentration`, `effective_number_of_members`, `largest_weight` | The Herfindahl index, the number of equal members it corresponds to, and the biggest weight | none |
| method | `prices(...)` | Candles for the basket as a whole, which every analysis method reads | `POST` `/prices` |
| method | `member_prices(...)`, `member_closes(...)`, `member_returns(...)` | Every member's candles, closes or returns, lined up by time; closes and returns are a data frame whose first column is `datetime`, followed by one column per label | `POST` `/prices` |
| method | `covariance_matrix(...)`, `correlation_matrix(...)` | How every pair of members moves together | `POST` `/prices` |
| method | `risk_contributions(...)` | Each member's share of the basket's volatility | `POST` `/prices` |
| method | `diversification_ratio(...)` | How much the members' moves cancel out | `POST` `/prices` |
| method | `return_contributions(...)` | What each member added to the basket's return | `POST` `/prices` |
| method | `top_gainers(count)`, `top_losers(count)` | The biggest movers today | `POST` `/ohlc` |
| method | `overlap_with(other)` | The weight two baskets have in common, the standard test of whether two funds differ | none |
| method | `add_member(member)`, `remove_member(instrument)` | Change the members in memory | none |

Every basket also inherits the fourteen analysis classes an instrument does, chained in R onto the shared `PriceAnalysis` base, the last of which is the [performance measures](analysis-performance.html), because its `prices()` method gives candles for the whole basket. Each candle is the sum over members of a fixed quantity times the member's candle. A weighted basket takes those quantities from its weights at the first candle of the range, starting from 100, which is how a price index moves between rebalances. The open and close are exact. The high and low are an approximation, because the members do not all reach their highs at the same moment, and volume and open interest are left empty.

#### Example

The example below builds a three-stock basket with one details request, then reads its live state and its history.

```{r}
rows <- list(
  list(
    exchange = "nse",
    segment = "equities",
    symbol = "INFY",
    weight = 50
  ),
  list(
    exchange = "nse",
    segment = "equities",
    symbol = "TCS",
    weight = 30
  ),
  list(
    exchange = "nse",
    segment = "equities",
    symbol = "WIPRO",
    weight = 20
  )
)
members <- MemberResolver$new()$resolve(rows)
basket <- AssetBasket$new("IT test", members)
print(basket$weights)
print(basket$breadth)
print(round(basket$risk_contributions(days = 365), 4))
print(round(basket$return_contributions(days = 365), 4))
print(basket$cumulative_return(days = 365))
```

When the Python version of this example ran on 2026-09-28, the weights came out as 0.5, 0.3 and 0.2, with one advancer and two decliners. Each member's contribution was negative: WIPRO −0.0631, TCS −0.0843 and INFY −0.1531. The three contributions add up to the basket's cumulative return of −30.06 percent, and INFY carried 56 percent of the risk on 50 percent of the weight.

#### Errors

The table below lists the conditions the constructor and the candle methods can signal.

| Condition | When |
|---|---|
| [`BasketMemberError`](guide-errors.html#basketmembererror) | The members are empty, name an instrument twice, or give weights to only some members; or UBI answered an error for a member when candles were asked for |
| `ValueError` | `unmapped_weight` is not between 0 and 1 |

## Portfolio

The constructor of [`Portfolio`](../reference/Portfolio.html) has the signature below.

```r
Portfolio$new(name, members, unified_broker_interface = NULL)
```

A portfolio gives every member a quantity, negative for a short position. Its weights are each member's share of today's value, and its candles show what the same quantities were worth at each candle, so its Sharpe ratio or drawdown describes the portfolio as you hold it now. The table below lists what a portfolio adds to every basket. The two builders are functions stored on the class generator, which is how the package writes Python's class methods.

| Kind | Member | Description | UBI route |
|---|---|---|---|
| function on the generator | `Portfolio$from_holdings(name = "holdings")` | Builds a portfolio of the account's long-term holdings | `GET` `/api/portfolio/holdings` |
| function on the generator | `Portfolio$from_positions(name = "positions", day = FALSE)` | Builds a portfolio of the open positions | `GET` `/api/portfolio/positions` |
| active binding | `quantities`, `values`, `value` | What is held and what it is worth at last prices | `POST` `/ltp` |
| active binding | `invested_value`, `unrealized_pnl` | What was paid and the profit on it | `POST` `/ltp` |
| active binding | `day_pnl`, `day_change_percent` | The profit and move since the previous close | `POST` `/ohlc` |
| method | `rebalance_trades(target, capital = NULL)` | The buys and sells that would give the portfolio another basket's weights, without sending anything | `POST` `/ltp` |
| places orders | `place_orders(product, transaction_type = "buy", ...)` | One market order per member, all in one request | `POST` `/api/orders/place` |
| places orders | `rebalance(target, product, ...)` | Sends the trades `rebalance_trades()` works out | `POST` `/api/orders/place` |

The orders go in UBI's list form of `POST /api/orders/place`, which takes up to 500 orders and places them in parallel, each at the broker that suits it. UBI's `basket` synthetic order is not used, because it takes at most 25 legs and sends every leg to the first leg's broker. Quantities are rounded towards zero to whole units and sent as computed, with no lot or tick check, because UBI checks orders itself.

#### Example

The example below reads the account's holdings as a portfolio.

```{r}
held <- Portfolio$from_holdings()
print(held)
print(held$value)
print(held$invested_value)
print(held$unrealized_pnl)
print(held$day_pnl)
```

When the Python version of this example ran on 2026-09-28, the portfolio held seven members, worth 9,240.43 rupees against 11,216.62 invested, an unrealised loss of 1,976.19 rupees and a loss of 216.29 rupees on the day. UBI's own holdings summary said 9,237.33 for the value at the same moment, because it prices each holding from its holdings document rather than from the live last price.

## Watchlist

The constructor of [`Watchlist`](../reference/Watchlist.html) has the signature below.

```r
Watchlist$new(name, instruments, unified_broker_interface = NULL)
```

A watchlist takes a list of plain instruments and weights them equally. It adds `add(instrument)` and `rank_by(column = "change_percent", ascending = FALSE)`, which sorts the members by any column of `ohlc` and signals `KeyError` for a column `ohlc` does not have. The example below ranks five shares by their move today.

```{r}
followed <- Watchlist$new(
  "banks and more",
  list(
    hdfc_bank,
    reliance,
    icici_bank,
    infosys,
    airtel
  )
)
ranked <- followed$rank_by("change_percent")
print(ranked[, c(
  "label",
  "last_price",
  "change_percent"
)])
```

When the Python version of this example ran, INFY led with a rise of 0.30 percent and the other four were down, from BHARTIARTL at −0.78 percent to RELIANCE at −2.32 percent.

## Index

The constructor of [`Index`](../reference/Index.html) has the signature below.

```r
Index$new(name, members, weighting = "stated", base_value = 100, base_date = NULL, linked_instrument = NULL, unified_broker_interface = NULL)
```

An index weights its members in one of three ways, which the table below lists.

| `weighting` | Each member's weight | Example |
|---|---|---|
| `stated` | Its own `weight`, as a factsheet gives it | NIFTY, loaded with its published weights |
| `equal` | The same for every member | NIFTY50 Equal Weight, or a file without weights |
| `price` | Its share of the sum of last prices, which is holding one unit of each | The Dow Jones |

`linked_instrument` is the official index the basket describes, when there is one, so `index$tracking_error(benchmark = index$linked_instrument)` measures how well the stored weights reproduce it. The `level` active binding reports today's level measured from `base_value` at `base_date`, and `to_portfolio(capital)` turns the index into whole units of each member for a sum of money, ready for `place_orders()`.

## ExchangeTradedFundConstituents

The constructor of [`ExchangeTradedFundConstituents`](../reference/ExchangeTradedFundConstituents.html) has the signature below.

```r
ExchangeTradedFundConstituents$new(name, members, fund = NULL, indicative_net_asset_value = NULL, unmapped_weight = 0.0, unified_broker_interface = NULL)
```

This is what an ETF holds, linked to the ETF through `fund`. `tracking_difference(...)` is the fund's return minus its holdings' return over a range, which is mostly the fund's fees. The `premium_or_discount` active binding compares the fund's last price with its indicative net asset value, when the index row that carries it is given as `indicative_net_asset_value`.

## MutualFundConstituents

The constructor of [`MutualFundConstituents`](../reference/MutualFundConstituents.html) has the signature below.

```r
MutualFundConstituents$new(name, members, fund = NULL, unmapped_weight = 0.0, unified_broker_interface = NULL)
```

This is what a mutual fund holds, linked to the scheme through `fund`. The `estimated_day_change_percent` active binding is the holdings' day move scaled down by `unmapped_weight`, the usual guess at today's change in the net asset value before the fund publishes it, and `estimated_net_asset_value(previous_net_asset_value)` applies it to the last published value. The weights come from the fund's monthly disclosure, so the basket is an estimate of the fund, not the fund.

## BasketStore and BasketCsvImporter

The constructor of [`BasketStore`](../reference/BasketStore.html) has the signature below.

```r
BasketStore$new(project_configuration = NULL, unified_broker_interface = NULL, collection = NULL)
```

The store saves and loads baskets in MongoDB. The third argument, `collection`, has no Python counterpart: it takes any object with the mongolite collection methods the store uses, so that the package's tests can pass an in-memory stand-in and never reach a database. Leave it as `NULL` in ordinary use, and the store connects to MongoDB on its first read or write. There is no `save()` method on a basket itself, so a basket is stored with `BasketStore$new()$save(basket)`. The table below lists the store's members.

| Kind | Member | Description |
|---|---|---|
| method | `save(basket, effective_date = NULL, source = "user")` | Stores the basket as the version in effect from a date, replacing one stored for the same date |
| method | `load(name, as_of = NULL)` | Rebuilds the version in effect on a day, as the class its stored kind names |
| method | `load_for_instrument(instrument, as_of = NULL)` | Finds the basket linked to an instrument; `constituents` calls this |
| method | `names(kind = NULL)`, `history(name)`, `delete(name, effective_date)` | List, inspect and remove stored versions |
| method | `build(document, linked_instrument = NULL)` | Builds a basket from a document without storing it |

`load()` signals [`BasketNotFoundError`](guide-errors.html#basketnotfounderror) when no version of the name is in effect on the day asked for, while `load_for_instrument()` returns `NULL` instead. A CSV file the importer cannot use signals [`BasketCsvImportError`](guide-errors.html#basketcsvimporterror), and every basket error is listed on [Errors](guide-errors.html#the-asset-basket-errors).

The constructor of [`BasketCsvImporter`](../reference/BasketCsvImporter.html) has the signature below; it passes `collection` through to the store in the same way.

```r
BasketCsvImporter$new(project_configuration = NULL, unified_broker_interface = NULL, collection = NULL)
```

The importer's one method, `import_file(path, name, kind = "index", exchange = "nse", segment = "equities", linked_instrument = NULL, effective_date = NULL, unmapped_weight = 0.0, source = "csv")`, reads a CSV with a `symbol` column and optional `exchange`, `segment`, `weight`, `quantity` and `instrument_id` columns. Column names are read without regard to case, so the NSE's own constituent files, whose header has `Symbol`, import as they are. A file without weights makes an equally weighted index. Weights may be fractions or percentages, with or without a `%` sign. The `kind` names the class the basket is stored and rebuilt as: `basket`, `portfolio`, `watchlist`, `index`, `exchange_traded_fund_constituents` or `mutual_fund_constituents`.

#### Example

The example below imports five NIFTY stocks with their weights, links them to the NIFTY index, and reads them back through `constituents`. The file `five.csv` holds the lines below.

```text
Company Name,Symbol,Weight
HDFC Bank,HDFCBANK,13.1%
Reliance Industries,RELIANCE,8.9%
ICICI Bank,ICICIBANK,8.5%
Infosys,INFY,5.2%
Bharti Airtel,BHARTIARTL,4.6%
```

The R code below imports it and reads it back.

```{r}
nifty <- EquityIndex$new(exchange = "nse", symbol = "NIFTY")
BasketCsvImporter$new()$import_file(
  "five.csv",
  name = "NIFTY FIVE",
  kind = "index",
  linked_instrument = nifty,
  effective_date = "2026-09-01"
)
five <- nifty$constituents
print(five)
print(round(five$weights, 4))
print(five$tracking_error(benchmark = nifty, days = 365))
```

When the Python version of this example ran on 2026-09-28, the basket came back as an `Index` named `NIFTY FIVE` with five members, whose weights, normalised to sum to 1, were 0.3251 for HDFCBANK, 0.2208 for RELIANCE, 0.2109 for ICICIBANK, 0.129 for INFY and 0.1141 for BHARTIARTL. Its tracking error against the NIFTY index over the year was about 0.060.

::: {.alert .alert-info}
**Under the hood.**

- Every live member sends one `POST` to UBI's list form of the instrument routes, described in [Several instruments at once](https://pramodathani.github.io/unified_broker_interface/rest-api/instruments/#several-instruments-at-once). UBI answers one entry per instrument, and a bad instrument fails only its own entry.
- `MemberResolver` in `R/asset_baskets_member_resolver.R` turns rows, each a named list, into members with one `POST /api/instruments/details`, and builds each instrument from its entry through the `details` argument of the `TradeableInstrument` and `NonTradeableInstrument` constructors, so no instrument is looked up twice.
- The reasoning behind each file is in `.claude/notes/R/asset_baskets_*.R.md`.
:::
